---
title: "Analysis of Engagement in Social Psychology and Biopsychology Courses"
author: "Ben Hicks"
output: html_document
---

The following data is from the Social Psychology (PSY203 and PSY453) and Biopsychology (PSY208 and PSY408) subjects run in 2018 Session 2. The aim of this report is to look at student engagement in these courses through the data available in Blackboard Learn. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
#knitr::opts_chunk$set(dev = 'png') - - uncomment to reduce file size
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(GGally))
suppressPackageStartupMessages(library(network))
suppressPackageStartupMessages(library(tidygraph))
suppressPackageStartupMessages(library(ggraph))
suppressPackageStartupMessages(library(ggthemes))
suppressPackageStartupMessages(library(lakit))
```

``` {r importing}
source(file.path('R','psych_functions.R'))
DATDIR <- file.path("~", "Data", "Psychology")
BIOAAFILE <- "aa Biopsychology 201860.csv"
SOCAAFILE <- "aa Social Psychology 201860.csv"
MARKSFILE <- "marks Psychology 201860.csv"
BIOCCFILE <- "cc Biopsychology 201860.csv"
SOCCCFILE <- "cc Social Psychology 201860.csv"
POSTSFILE <- "posts Psychology 201860.csv"


# Activity Accumulator Data - - - - - - - - - - - - - 
tidyAA <- function(df) {
  df$id <- factor(df$id)
  df <- df %>%
    filter(!is.na(id)) %>%
    mutate(forum_bool = str_detect(tolower(data), "discus|thread|forum"),
                                      content = tidyDataField(data),
                                      name = paste0(lastname, ', ', firstname)) %>%
    rename(session = session_id)
  df <- df %>% mutate(forum_bool = if_else(is.na(forum_bool), FALSE, forum_bool))
  return(df)
} # -  - - - - - - --- - - - - - --- - - - --- - - -- 
socpsyAct18 <- read_csv(file.path(DATDIR, SOCAAFILE)) 
socpsyAct18 <- tidyAA(socpsyAct18)

biopsyAct18 <- read_csv(file.path(DATDIR, BIOAAFILE))
biopsyAct18 <- tidyAA(biopsyAct18)

# Node preparation (i.e. data for each student)

# Mark data
marks18 <- read_csv(file.path(DATDIR, MARKSFILE)) %>%
  mutate(mark = score / possible) 
marksSummary18 <- marks18 %>% 
  group_by(subject_name, student, id) %>%
  summarise(avg_grade = mean(mark, na.rm = T))
marksSoc18 <- marks18 %>% filter(subject_name == "Social Psychology")
marksBio18 <- marks18 %>% filter(subject_name == "Biopsychology")

# Discussion views from aa ------------------------------------------
discussionViews <- function(df) {
  df %>%
    select(name, subject, forum_bool, data) %>%
    filter(  forum_bool |
           str_detect(tolower(data), "discus|thread|forum")
         ) %>% 
    group_by(subject, name) %>%
    summarise(ForumViews = n()) %>%
    group_by(name) %>%
    summarise(ForumViews = sum(ForumViews)) %>%
    rename(Name = name)  
} # ------------------------------------------------------------------
discussionViewsBio <- discussionViews(biopsyAct18)
discussionViewsSoc <- discussionViews(socpsyAct18)

# Other interaction data from aa
aaSummaryBio <- aa_summary_by_user(biopsyAct18)
aaSummarySoc <- aa_summary_by_user(socpsyAct18)

# Node assembly

studentNodeData <- studentNodeData %>% plyr::rename(replace = c("Avg Learn Grade" = "Grade",
                                                  "Course item Interactions" = "Interactions",
                                                  "Course Accesses" = "Accesses"))

teacherNodeData <- roleData %>% filter(ID %in% setdiff(roleData$ID, studentNodeData$ID) & ID %in% snData$MsgFrom)
teacherNodeData$Grade <- NA
teacherNodeData$Interactions <- NA
teacherNodeData$Accesses <- 1
nodeData <- rbind(studentNodeData, teacherNodeData)

# Building network
edgesAll <- data.frame(from = snData$MsgFrom, 
                       to = snData$MsgTo, 
                       weight = snData$Strength, 
                       Subject = snData$Subject)
questions <- edgesAll %>% 
  filter(to == "NULL") %>% 
  group_by(from, Subject) %>% 
  summarise(Questions = sum(weight))
edges <- edgesAll %>% filter(to != "NULL")

nodes <- data.frame(ID = as.factor(nodeData$ID), 
                Name = as.character(nodeData$Name), 
                Role = nodeData$Role,
                Accesses = nodeData$Accesses,
                Interactions = nodeData$Interactions,
                Grade = nodeData$Grade,
                Subject = nodeData$Subject
)


missingNodeDataIDs <- setdiff(union(edges$from, edges$to), nodes$ID)

# Editing subject field - might be redundant
nodes <- nodes %>% 
  mutate(Subject = ifelse(str_detect(Subject, "208|458"), "Biopsychology", "Social Psychology")) %>%
  unique() 
edges <- edges %>% 
  mutate(Subject = ifelse(str_detect(Subject, "208|458"), "Biopsychology", "Social Psychology")) %>%
  unique()



# Graphing in tidygraph and ggraph

nodesBio <- nodes %>% filter(str_detect(Subject, "Bio"))
edgesBio <- edges %>% filter(str_detect(Subject, "Bio"))
nodesSoc <- nodes %>% filter(str_detect(Subject, "Social"))
edgesSoc <- edges %>% filter(str_detect(Subject, "Social"))

# Filtering strange IDs
#nodesBio <- nodesBio %>% filter(ID %in% discussionViewsBio$ID | !(Name %in% discussionViewsBio$Name))
#nodesSoc <- nodesSoc %>% filter(ID %in% discussionViewsSoc$ID | !(Name %in% discussionViewsSoc$Name))

# getting grade quartiles
nodesBio <- nodesBio %>% mutate(grade_quartile = ntile(Grade, 4)) 
#nodesBio[is.na(nodesBio$grade_quartile),]$grade_quartile <- 0

nodesSoc <- nodesSoc %>% mutate(grade_quartile = ntile(Grade, 4))
#nodesSoc[is.na(nodesSoc$grade_quartile),]$grade_quartile <- 0


# merging with discussion data
nodesBio <- merge(nodesBio, discussionViewsBio, all.x = T)
nodesBio[is.na(nodesBio$ForumViews),]$ForumViews <- 0
nodesSoc <- merge(nodesSoc, discussionViewsSoc, all.x = T)
nodesSoc[is.na(nodesSoc$ForumViews),]$ForumViews <- 0

# converting NAs in Accesses to 0
nodesBio <- nodesBio %>% mutate(Accesses = if_else(is.na(Accesses), 0, Accesses))
nodesSoc <- nodesSoc %>% mutate(Accesses = if_else(is.na(Accesses), 0, Accesses))

# Making graph structures
BioPsyNet <- tbl_graph(nodes = nodesBio, edges = edgesBio) 
SocPsyNet <- tbl_graph(nodes = nodesSoc, edges = edgesSoc)

```

# Activity

## Timeline Averages

``` {r timelines}
timelineData <- timelineDataRaw %>% mutate(subject = substr(`subject code`, 3, 8),
                                           session = as.factor(substr(`subject code`, 10, 15)))
timeSum <- timelineData[timelineData$subject != "PSY458",] %>% 
  group_by(week, measure, subject, session) %>%
  summarise(score = sum(score, na.rm = T))

# collecting together different campuses and ommitting some data
nonDiscussion <- c("Avg Site Accesses by Week of Term","Avg Interactions per Week","Avg Minutes per Site Access","Avg Minutes by Day")
timeSumShort <- timelineData[timelineData$subject != "PSY458" &
                              timelineData$measure %in% nonDiscussion,] %>% 
  group_by(week, measure, subject, session) %>%
  summarise(score = sum(score, na.rm = T))
timePlot <- ggplot(timeSumShort, aes(x = week, y = score, color = session)) +
  geom_line() +
  facet_grid(measure ~ subject, scales = "free") +
  theme_tufte() +
  theme(strip.text.y = element_text(size = 8, angle = 0),
        axis.label.y = NULL) +
  scale_color_brewer(palette = "Paired")
timePlot + ggtitle("Average Course Activity by Week of Course")

# t-testing 1st and 3rd graphs for PSY203
t1int <- timeSumShort[timeSumShort$measure == "Avg Interactions per Week" & timeSumShort$subject == "PSY203" & timeSumShort$session == "201760",]$score
t2int <- timeSumShort[timeSumShort$measure == "Avg Interactions per Week" & timeSumShort$subject == "PSY203" & timeSumShort$session == "201860",]$score
t1Min <- timeSumShort[timeSumShort$measure == "Avg Minutes per Site Access" & timeSumShort$subject == "PSY203" & timeSumShort$session == "201760",]$score
t2Min <- timeSumShort[timeSumShort$measure == "Avg Minutes per Site Access" & timeSumShort$subject == "PSY203" & timeSumShort$session == "201860",]$score

ttestInt <- t.test(t2int, t1int, paired = T)
ttestMin <- t.test(t2Min, t1Min, paired = T)
```

The most significant difference between sessions of the results above is for Social Psychology (PSY203) on the *Average Interactions per Week* and *Avg Minutes per Site Access* metrics. The interactions show an increase of between `r round(ttestInt$conf.int[[1]],2)` and `r round(ttestInt$conf.int[[2]], 2)` (95% confidence, paired t-test), and a similar test on the minutes per site access shows an increase of between `r round(ttestMin$conf.int[[1]],2)` and `r round(ttestMin$conf.int[[2]], 2)`. For the latter, it is worth flagging that mean duration on site can be thrown by students leaving the browser open - median would be a better measure here.

## Item Activity

We can examine student interactions with the course content at a finer level, and filter according to activities of note.


### Social Psychology

``` {r social_psych_usage}
# Using the ACTIVITY_ACCUMULATOR data to examine item activity


socUsage <- socpsyAct18 %>%
  select(name, timestamp, content, role, forum_bool) %>%
  filter(role == "S" & timestamp < as.Date('2018-11-01')) %>%
  mutate(Activity = if_else(
    str_detect(content, "Week 2: The Social Self|Week 3:Perc|Week 4:Stereo|Week 5:Att|Week 6:Conf|Week 10:Att|Week 11:Help|Applied Social"), 
    "Tutorial",
    if_else(forum_bool, "Forum", "Other")
    )
    )%>%
  mutate(Activity = if_else(is.na(Activity), "Other", Activity)) %>%
  plyr::rename(replace = c("name" = "Name")) %>% # ERROR IN THIS NEXT LINE, JOIN NOT RECOGNISING AND RETURNING NA
  left_join(nodesSoc %>% 
          select(Name, Accesses, Interactions, grade_quartile, ForumViews) %>% 
          unique(), by = c("Name"))

socGameActPlot <- socUsage %>% 
  ggplot(aes(x = timestamp, y = grade_quartile, color = ForumViews)) +
  scale_color_gradient2(name = "Forum Participation", 
                         low = "Black",
                          mid = "Dark Blue",
                         high = "Red",
                         midpoint = 3.5,
                         trans = "log1p",
                         breaks = c(2, 10, 50, 250, 1250)) +  
  geom_jitter(alpha = 0.01) +
  scale_y_continuous(name = "Grade Quartile") +
  ggthemes::theme_tufte() + 
  theme(axis.title.x = element_blank()) + 
  facet_grid(Activity ~ .) +
  ggtitle("Usage of items in Social Psychology")

socUsage %>% group_by(Activity) %>% tally() %>% knitr::kable()

socGameActPlot

```

``` {r eval = T}
# Looking at what was clicked prior to forum access

socForumExamine <- socpsyAct18 %>% filter(role == "S") %>% select(id, timestamp, content, forum_bool) %>% arrange(id, timestamp)
r = length(socForumExamine$id)
socForumAccess <- data.frame(from = socForumExamine[1:r - 1, "content"],
                             to = socForumExamine[2:r, "content"],
                             idFrom = socForumExamine[1:r - 1, "id"],
                             idTo = socForumExamine[2:r, "id"],
                             toForum = socForumExamine[2:r, "forum_bool"])
names(socForumAccess) <- c("from", "to", "idFrom", "idTo", "toForum")
socForumAccess <- socForumAccess %>% filter(from != to, idFrom == idTo) # was filtering for forum true as well.
socForumAccessEdges <- socForumAccess %>% 
  filter(from != to, idFrom == idTo, toForum) %>% 
  select(from, to) %>% 
  group_by(from, to) %>% 
  summarise(weight = n())

socForumAccessNodes <- socForumExamine %>% 
  select(content, forum_bool, timestamp) %>%
  group_by(content, forum_bool) %>%
  summarise(first_access = min(timestamp), hits = n()) %>%
  mutate(PageType = if_else(str_detect(content, "Week 2: The Social Self|Week 3:Perc|Week 4:Stereo|Week 5:Att|Week 6:Conf|Week 10:Att|Week 11:Help|Applied Social"), "Tutorial", 
                            if_else(forum_bool, "Forum", "Other"))) %>%
  mutate(PageType = if_else(is.na(PageType), "Other", PageType)) %>%
  arrange(first_access)

socForumAccessGraph <- tbl_graph(nodes = socForumAccessNodes,
                                 edges = socForumAccessEdges,
                                 directed = TRUE)

socFAplot <- socForumAccessGraph %>% activate(nodes) %>% filter(!node_is_isolated()) %>% 
  ggraph(layout = "linear", circular = T) + 
  geom_node_point(aes(size = hits, color = PageType)) +
  geom_edge_arc2(aes(alpha = weight)) +
  scale_edge_alpha(guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_color_brewer(palette = "Dark2") +
  geom_node_text(aes(label = content, alpha = hits/3), size = 2,repel = T) +
  scale_alpha(guide = "none") +
  theme_graph() 

socFAplot + ggtitle("Social Psychology Site Navigation")
```

\newpage

### Biopsychology Activity

``` {r biopsych_usage}
# Using the ACTIVITY_ACCUMULATOR data to examine item activity

bioUsage <- biopsyAct18 %>%
  select(name, timestamp, content, role, forum_bool) %>%
  filter(role == "S" ) %>%
  #& timestamp < as.Date('2018-11-01')) %>%
  mutate(Activity = if_else(str_detect(tolower(content), "game"), "Game",
                                   if_else(str_detect(tolower(content), "in the lab"), "In the lab",
                                           if_else(forum_bool, "Forum", "Other")))) %>%
  mutate(Activity = if_else(is.na(Activity), "Other", Activity)) %>%
  plyr::rename(replace = c("name" = "Name")) %>%
  merge(nodesBio %>% select(Name, Accesses, Interactions, grade_quartile, ForumViews), all.x = T)



bioGameActPlot <- bioUsage %>% 
  ggplot(aes(x = timestamp, y = grade_quartile, color = ForumViews)) +
  scale_color_gradient2(name = "Forum Participation", 
                         low = "Black",
                          mid = "Dark Blue",
                         high = "Red",
                         midpoint = 3.5,
                         trans = "log1p",
                         breaks = c(2, 10, 50, 250, 1250)) +  
  geom_jitter(alpha = 0.002) +
  scale_y_continuous(name = "Grade Quartile") +
  ggthemes::theme_tufte() + 
  theme(axis.title.x = element_blank()) + 
  facet_grid(Activity ~ .) +
  ggtitle("Usage of items in Biopsychology")

bioUsage %>% group_by(Activity) %>% tally() %>% knitr::kable()

bioGameActPlot

bioGameActPlotFocus <- bioUsage %>% filter(timestamp < as.Date('2018-10-15')) %>%
  ggplot(aes(x = timestamp, y = grade_quartile, color = ForumViews)) +
  scale_color_gradient2(name = "Forum Participation", 
                         low = "Black",
                          mid = "Dark Blue",
                         high = "Red",
                         midpoint = 3.5,
                         trans = "log1p",
                         breaks = c(2, 10, 50, 250, 1250)) +  
  geom_jitter(alpha = 0.002) +
  scale_y_continuous(name = "Grade Quartile") +
  ggthemes::theme_tufte() + 
  theme(axis.title.x = element_blank()) + 
  facet_grid(Activity ~ .) +
  ggtitle("Usage of items in Biopsychology - focus on active session")


bioGameActPlotFocus
```

``` {r Biopsych_forum_access, eval = T}
# Looking at what was clicked prior to forum access

bioForumExamine <- biopsyAct18 %>% filter(role == "S") %>% select(id, timestamp, content, forum_bool) %>% arrange(id, timestamp)
r = length(bioForumExamine$id)
bioForumAccess <- data.frame(from = bioForumExamine[1:r - 1, "content"],
                             to = bioForumExamine[2:r, "content"],
                             idFrom = bioForumExamine[1:r - 1, "id"],
                             idTo = bioForumExamine[2:r, "id"],
                             forum = bioForumExamine[2:r, "forum_bool"])
names(bioForumAccess) <- c("from", "to", "idFrom", "idTo", "toForum")

bioForumAccessEdges <- bioForumAccess %>% 
  filter(from != to, idFrom == idTo, toForum) %>% 
  select(from, to) %>% 
  group_by(from, to) %>% 
  summarise(weight = n())

bioForumAccessNodes <- bioForumExamine %>% 
  select(content, forum_bool, timestamp) %>%
  group_by(content, forum_bool) %>%
  summarise(first_access = min(timestamp), hits = n()) %>%
  mutate(PageType = if_else(str_detect(content, "lab|Lab"), "Lab", 
                            if_else(forum_bool, "Forum", "Other"))) %>%
  mutate(PageType = if_else(is.na(PageType), "Other", PageType)) %>%
  arrange(first_access)

bioForumAccessGraph <- tbl_graph(nodes = bioForumAccessNodes,
                                 edges = bioForumAccessEdges,
                                 directed = TRUE)

bioFAplot <- bioForumAccessGraph  %>% activate(nodes) %>% filter(!node_is_isolated()) %>% 
  ggraph(layout = "linear", circular = T) + 
  geom_node_point(aes(size = hits, color = PageType)) +
  geom_edge_arc2(aes(alpha = weight)) +
  scale_edge_alpha(guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_color_brewer(palette = "Dark2") +
  geom_node_text(aes(label = content, alpha = hits/3), size = 2,repel = T) +
  scale_alpha(guide = "none") +
  theme_graph() 

bioFAplot + ggtitle("Biopsychology Site Navigation")
```

\newpage

# Network analysis on forum posts and replies

Discussion (forum) posts were extraced, and converted into networks. The main algorithms used are:

- _infomap_ for clustering (forming communities)
- _central authority_ for importance
- the force-directed _Kamada Kawai_ algorithm for graph layout

Each *node* in the network represents a person in the course. Each *edge* is the sum of the posts between the people, darker lines mean more posts / replies. Grades are displayed in quartiles, with 4 being the upper quartile and represented by a larger disc.

## Social Psychology

``` {r fig.height = 4}
# Graphs for Social Psychology


gSocAll <- buildGraph(SocPsyNet)
gSocAll + ggtitle("All Posts for PSY203: Social Psychology") #+ geom_node_text(aes(label = id), size = 2)

gSocNet <- buildTrimGraph(SocPsyNet)
gSocNet + ggtitle("All Posts for PSY203: Social Psychology (isolated points removed)")

gSocLurker <- buildLurkerGraph(SocPsyNet %>% activate(nodes) %>% filter(Role == "S"))
gSocLurker + ggtitle("All Posts for PSY203: Social Psychology, with passive forum interactions")

gSocLurkerZoom <- buildLurkerGraph(SocPsyNet %>% activate(nodes) %>% filter(Role == "S" & !(node_is_isolated())))
gSocLurkerZoom + ggtitle("All Posts for PSY203: Social Psychology, with passive forum interactions (isolated points removed)")

gSocStudents <- buildTrimGraph(SocPsyNet %>% activate(nodes) %>% filter(Role == "S")) 
gSocStudents + ggtitle("Student only interactions for PSY203: Social Psychology (isolated points removed)")

gSocImportance <- buildImportanceGraph(SocPsyNet %>% activate(nodes) %>% filter(Role == "S"))
gSocImportance + ggtitle("Student only interaction by Importance for PSY203: Social Psychology (isolated points removed)")
  
gSocCommunity <- buildClusterGraph(SocPsyNet %>% activate(nodes) %>% filter(Role == "S"))
gSocCommunity + ggtitle("Student only Communities for PSY203: Social Psychology (isolated points removed)")

```

\newpage

## Biopsychology

``` {r fig.height = 4}

gBioAll <- buildGraph(BioPsyNet)
gBioAll + ggtitle("All Posts for PSY208: Biopsychology") #+ geom_node_text(aes(label = id), size = 2)

gBioNet <- buildTrimGraph(BioPsyNet)
gBioNet + ggtitle("All Posts for PSY208: Biopsychology (isolated points removed)")

gBioLurker <- buildLurkerGraph(BioPsyNet %>% activate(nodes) %>% filter(Role == "S"))
gBioLurker + ggtitle("All Posts for PSY208: Biopsychology, with passive forum interactions")

gBioLurkerZoom <- buildLurkerGraph(BioPsyNet %>% activate(nodes) %>% filter(Role == "S" & !(node_is_isolated())))
gBioLurkerZoom + ggtitle("All Posts for PSY208: Biopsychology, with passive forum interactions (isolated points removed)")

gBioStudents <- buildTrimGraph(BioPsyNet %>% activate(nodes) %>% filter(Role == "S")) 
gBioStudents + ggtitle("Student only interactions for PSY208: Biopsychology (isolated points removed)")

gBioImportance <- buildImportanceGraph(BioPsyNet %>% activate(nodes) %>% filter(Role == "S"))
gBioImportance + ggtitle("Student only interaction by Importance for PSY208: Biopsychology (isolated points removed)")
  
gBioCommunity <- buildClusterGraph(BioPsyNet %>% activate(nodes) %>% filter(Role == "S"))
gBioCommunity + ggtitle("Student only Communities for PSY208: Biopsychology (isolated points removed)")

```
