---
title: "Layering Network Graphs"
output: slidy_presentation
---

```{r setup}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, cache = TRUE)
```

``` {r loading_data}
# loading data
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggraph))
suppressPackageStartupMessages(library(tidygraph))
suppressPackageStartupMessages(library(gganimate))
DATDIR <- file.path("~", "Data", "Psychology")
BIOAAFILE <- "aa Biopsychology 201860.csv"
BIOCCFILE <- "cc Biopsychology 201860.csv"


aa_full <- read_csv(file.path(DATDIR, BIOAAFILE))
cc_full <- read_csv(file.path(DATDIR, BIOCCFILE)) 

aa <- aa_full %>% 
  filter(role == 'S', !is.na(content_pk1), timestamp < as.POSIXct('2018-10-30')) %>% 
  select(id, time = timestamp, content_pk1, session = session_id) %>% 
  mutate(id = factor(id), content_pk1 = factor(content_pk1)) %>% 
  arrange(id, time)

cc <- cc_full %>% 
  select(pk1, parent_pk1, title) %>% 
  mutate(pk1 = factor(pk1), parent_pk1 = factor(parent_pk1))

# creating edgelist from aa
aa_edges <- tibble(from = aa$content_pk1[1:(nrow(aa) - 1)],
                   to =   aa$content_pk1[2:nrow(aa)],
                   time = aa$time[2:nrow(aa)],
                   student_from = aa$id[1:(nrow(aa) - 1)],
                   student = aa$id[2:nrow(aa)],
                   session_from = aa$session[1:(nrow(aa) - 1)],
                   session = aa$session[2:nrow(aa)])
# filter out any pairs not belonging to the same id and session
aa_edges <- aa_edges %>% 
  filter(student == student_from, session == session_from) %>% 
  select(from, to, time, student) %>% 
  mutate(weight = 2)

cc_edges <- cc %>% 
  select(from = parent_pk1, to = pk1) %>% 
  filter(!is.na(from)) %>% 
  mutate(weight = 1, time = as.POSIXct('2018-05-01')) # used manual set time - needs to be well before other data

# creating node lists
cc_nodes <- cc %>% select(id = pk1, title)
aa_nodes <- aa %>% select(id = content_pk1) %>% unique()

cc_graph <- tbl_graph(nodes = cc_nodes, edges = cc_edges)
cc_layout <- create_layout(cc_graph, layout = "nicely")
np <- data.frame(x = cc_layout$x, y = cc_layout$y)

aa_graph <- tbl_graph(nodes = cc_nodes, edges = aa_edges)
aa_sg_3 <- aa_graph %>% activate(edges) %>% 
  filter(student %in% sample(student, 3))
aa_sg_1 <- aa_graph %>% activate(edges) %>% 
  filter(student %in% sample(student, 1))

joined_graph_1 <- cc_graph %>% graph_join(aa_sg_1) 
joined_graph_1 <- joined_graph_1 %>% 
  activate(edges) %>% 
  mutate(student = if_else(is.na(student), "_course structure", as.character(student)))

joined_graph_3 <- cc_graph %>% graph_join(aa_sg_3) 
joined_graph_3 <- joined_graph_3 %>% 
  activate(edges) %>% 
  mutate(student = if_else(is.na(student), "_course structure", as.character(student)))

```


## Base Network

Original network. This shows the connections between different parts of the subject site.


``` {r draw_base_graph}
g_0 <- ggraph(joined_graph_1 %>% 
                activate(edges) %>% 
                filter(student == '_course structure'), 
            layout = "manual", 
            node.positions = np) + 
  geom_node_point() + 
  geom_edge_link0() +
  theme_graph() +
  theme(legend.position = "none",
        text = element_text(family = "TT Times New Roman"))

g_0
```

## Network with a single student's activity

Black lines represent the connections between the site. Blue through to red lines represent the student navigating their path around the site.

```{r single_activity}
g_1 <- ggraph(joined_graph_1, layout = "manual", node.positions = np) + 
  geom_node_point() + 
  geom_edge_link0(aes(color = as.numeric(time), 
                      width = weight, 
                      alpha = 2 - weight,
                      group = time)) +
  scale_edge_alpha_continuous(range = c(0.1, 1)) +
  scale_edge_width(range = c(0.6,1)) +
  scale_edge_color_gradientn(colors = c("black", "#47ADED", "#D81C41")) +
  theme_graph() +
  theme(legend.position = "none",
        text = element_text(family = "TT Times New Roman"))

g_1
```

## Comparing several student's activity


```{r facet_graph}
g_3 <- ggraph(joined_graph_3, layout = "manual", node.positions = np) + 
  geom_node_point() + 
  geom_edge_link0(aes(color = as.numeric(time), 
                      width = weight, 
                      alpha = 2 - weight,
                      group = time)) +
  scale_edge_alpha_continuous(range = c(0.1, 1)) +
  scale_edge_width(range = c(0.6,1)) +
  scale_edge_color_gradientn(colors = c("black", "#47ADED", "#D81C41")) +
  theme_graph() +
  theme(legend.position = "none",
        text = element_text(family = "TT Times New Roman"),
        strip.text = element_blank())

g_facet <- g_3 +
  facet_edges(~ student)
g_facet
```

## Animating a single student's activity


``` {r animate_graph}
g_animate <- g_1 +
  transition_time(time) +
  shadow_mark(1, colour = "black") +
  labs(title = '{frame_time}')
g_animate
```



