---
title: "Hybrid Exploration of Student Activity in Online Environment"
output: html_document
---

<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 8px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
suppressPackageStartupMessages(library(lakit))
suppressPackageStartupMessages(library(igraph))
suppressPackageStartupMessages(library(tidygraph))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(ggraph))
suppressPackageStartupMessages(library(flextable))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(ggforce))
source(file.path('~', 'csu-psychology-engagement', 'R', 'psych_functions.R'))

# Loading data
aabio <- rdata.psych::aaBio18 %>% as_tibble()
aasoc <- rdata.psych::aaSoc18 %>% as_tibble()
ccbio <- rdata.psych::ccBio18 %>% as_tibble()
ccsoc <- rdata.psych::ccSoc18 %>% as_tibble()
nodesBio <- rdata.psych::nodesBio18 %>% 
  filter(Role == "S") %>% 
  as_tibble()
nodesSoc <- rdata.psych::nodesSoc18 %>% 
  filter(Role == "S") %>% 
  as_tibble()

# Classifying cc
ccbio$type <- ccbio$title %>% map_chr(classify_bio_title)
ccsoc$type <- ccsoc$title %>% map_chr(classify_soc_title)

# Imputing NA's in content
aabio$content_pk1 <- impute_na_with_previous(aabio$content_pk1)
aasoc$content_pk1 <- impute_na_with_previous(aasoc$content_pk1)

# Joining to get 'title' field
aabio <- left_join(aabio, ccbio %>% select(title, type, content_pk1 = pk1))
aasoc <- left_join(aasoc, ccsoc %>% select(title, type, content_pk1 = pk1))

ZERODATE <- '2018-06-04'
SESSIONSTART <- '2018-07-09'
SESSIONEND <- '2018-10-26'

# TODO: Classify bio and soc content, then feed into 
# activity spectrums
classify_bio_content <- function(content){
  # Do stuff
}

# Functions
make_cc_graph <- function(cc){
  nds <- tibble(id = as.character(cc$pk1),
                title = cc$title, 
                handle = cc$handle)
  eds <- tibble(from = as.character(cc$parent_pk1), 
                to = as.character(cc$pk1), 
                alpha = 1,
                timestamp = as.POSIXct(ZERODATE),
                cluster = 0) %>% 
    drop_na()
  return(tbl_graph(nodes = nds, edges = eds))
}

make_aa_graph <- function(aa, compress = TRUE, compress_by = c("weeks", "months")) {
  n <- length(unique(aa$person))
  df <- aa %>% 
    select(person, session, timestamp, content_pk1, cluster) %>% 
    arrange(person, timestamp) %>% 
    drop_na() %>% 
    filter(timestamp <= SESSIONEND, timestamp >= SESSIONSTART)
  end <- nrow(df)
  ids <- unique(df$content_pk1)
  nds <- tibble(id = as.character(ids))
  eds <- tibble(from = as.character(df$content_pk1[1:(end - 1)]),
                to =   as.character(df$content_pk1[2:end]),
                person_from = df$person[1:(end - 1)],
                person_to = df$person[2:end],
                session_from = df$session[1:(end - 1)],
                session_to = df$session[2:end],
                timestamp = df$timestamp[2:end],
                cluster = df$cluster[1:(end - 1)],
                alpha = 1/log(n+1, base = 2)) %>% # Inverse information scale 
    filter(person_from == person_to, session_from == session_to) %>% 
    select(-person_from, -session_from, -session_to) %>% 
    rename(person = person_to)
  if (compress) {
    if (compress_by == "weeks") {
      eds$week <- week(eds$timestamp) - week(SESSIONSTART) + 1
      eds <- eds %>% 
        group_by(from, to, week, cluster) %>% 
        summarise(alpha = sum(alpha/n, na.rm = TRUE))
    } else if (compress_by == "months") {
      eds$month <- month(eds$timestamp) - month(SESSIONSTART)
      eds <- eds %>% 
        group_by(from, to, month, cluster) %>% 
        summarise(alpha = sum(alpha/n, na.rm = TRUE))
    } else {
      break("Unrecognised compress_by variable, must be 'weeks' or 'months'")
    }
  } else {
    eds <- eds %>%
      group_by(from, to, timestamp, cluster) %>%
      summarise(alpha = sum(alpha/n, na.rm = TRUE))
  }
  return(tbl_graph(nodes = nds, edges = eds))
  
}

draw_cc <- function(graph, layout) {
  ggraph(graph, layout = "manual", 
         x = layout$x, y = layout$y) +
    geom_node_point(aes(color = handle)) +
    geom_node_text(aes(label = title), alpha = 0.8, size = 1.5) +
    geom_edge_link() +
    theme_graph() +
    theme(legend.position = "none")
}

draw_ag <- function(graph, layout) {
  ggraph(graph , layout = "manual", 
         x = layout$x, y = layout$y) +
  # ,layout = "nicely") +
    geom_node_point(size = 0.5, alpha = 0.3) + # removed 'handle' as color aes
    geom_edge_link(aes(alpha = alpha)) +
    theme_graph() + 
    theme(legend.position = "none",
          strip.background = element_blank(),
          strip.text.x = element_blank()) 
}

draw_ag_timeColor_week <- function(graph, layout) {
  nodes <- graph %>% 
    activate(nodes) %>% 
    as_tibble() %>% 
    inner_join(layout %>% select(id, x, y), by = "id")
  graph <- graph %>% 
    activate(nodes) %>% 
    filter(id %in% nodes$id)
  ggraph(graph, layout = "manual", 
         x = nodes$x, y = nodes$y) +
    geom_edge_link(aes(alpha = alpha*0.4, color = week)) +
    geom_node_point(size = 2, alpha = 0.8) +
    scale_edge_color_continuous(low = "blue", high = "dark red") +
    scale_color_brewer(palette = "Set2") +
    theme_graph() + 
    theme(legend.position = "none",
          strip.background = element_blank()) 
}

draw_ag_timeColor(ag_m, lyt) + facet_grid(month ~ cluster)
```

# Course Structure

The subject sites on Blackboard have a built in hierarchy; content on the site has a folder structure, so each content item may have a single *parent* which it is contained within.This leads to a natural way of displaying the course structure, with the *parent* to *child* relationships forming the edges of the network.

``` {r setting_up_graphs}
set.seed(18)

lyt <- create_layout(cg, layout = "nicely")
draw_cc(cc, lyt)
draw_ag(ag_m, lyt) + facet_grid(month ~ cluster)

cg_bio <- make_cc_graph(ccbio)
bio_layout <- create_layout(cg_bio, layout = "nicely")
bio_clusters <- min(nodesBio$cluster_5, na.rm = T):max(nodesBio$cluster_5, na.rm = T)
# bio_act_by_cluster <- bio_clusters %>% 
#   map(function(n){aabio %>% 
#       rename(cluster = cluster_2) %>% 
#       make_aa_graph(compress_by = "weeks")})
bio_act_by_cluster <- make_aa_graph(aabio %>% 
                                      rename(cluster = cluster_5) %>% 
                                      filter(cluster > 0), 
                                    compress_by = "weeks")
jg_clusters_bio <- cg_bio
for (n in bio_clusters) {
  jg_clusters_bio <- jg_clusters_bio %>% graph_join(bio_act_by_cluster[[n]], by = "id")
}


bio_cc_plot <- draw_cc(cg_bio, bio_layout)

bio_cluster_act_plot <- draw_ag_timeColor(bio_act_by_cluster,
                                bio_layout)

set.seed(5)  
cg_soc <- make_cc_graph(ccsoc)
soc_layout <- create_layout(cg_soc, layout = "nicely")
# ag_1_soc <- aasoc %>% filter(cluster == 1) %>% 
#   make_aa_graph()
# ag_2_soc <- aasoc %>% filter(cluster == 2) %>% 
#   make_aa_graph()
# jg_12_soc <- cg_soc %>% graph_join(ag_1_soc) %>% graph_join(ag_2_soc)

soc_clusters <- min(nodesSoc$cluster, na.rm = T):max(nodesSoc$cluster, na.rm = T)
soc_act_by_cluster <- soc_clusters %>% 
  map(function(n){aasoc %>% filter(cluster == n) %>% make_aa_graph(date_floor = "weeks")})
jg_clusters_soc <- cg_soc
for (n in soc_clusters) {
  jg_clusters_soc <- jg_clusters_soc %>% graph_join(soc_act_by_cluster[[n - 2]], by = "id")
}


soc_cc_plot <- draw_cc(cg_soc, soc_layout)
  
# soc_12act_plot <- draw_ag(jg_12_soc %>%   
#                             activate(edges) %>%
#                             filter(cluster > 0) %>%
#                             activate(nodes) %>%
#                             filter(id %in% soc_layout$id)
#                           , soc_layout)

ag_soc <- make_aa_graph(aasoc %>% rename(cluster = cluster_3), compress_by = "weeks")
soc_cluster_act_plot <- draw_ag(ag_soc %>% 
                                 activate(edges) %>% 
                                 filter(cluster > 0) %>% 
                                 activate(nodes),
                                soc_layout)

soc_cluster_act_plot_timeColor <- draw_ag_timeColor_week(ag_soc %>% 
                                 activate(edges) %>% 
                                 filter(cluster > 0) %>% 
                                 activate(nodes) %>% 
                                 filter(id %in% soc_layout$id),
                                soc_layout) +
  theme(strip.text = element_blank())

sml_lyt <- soc_layout %>% 
  filter(id %in% ag_soc())

jg_soc <- make_cc_graph(ccsoc)
lyt <- create_layout(jg_soc, layout = "nicely")
jg_soc <- graph_join(jg_soc, make_aa_graph(aasoc %>% 
                                             rename(cluster = cluster_3) %>% 
                                             filter(content_pk1 %in% ccsoc$pk1), 
                                           compress_by = "weeks"), 
                     by = "id")

soc_for_alasi <- jg_soc %>% 
  ggraph(layout = "manual", x = lyt$x, y = lyt$y) +
  geom_edge_link(aes(alpha = alpha*0.2, color = week))+
    scale_edge_color_continuous(low = "blue", high = "dark red") +
    theme_graph() + 
    theme(legend.position = "none")

soc_for_alasi <- ag_soc %>% 
  activate(edges) %>% 
  filter(cluster > 0) %>% 
  activate(nodes) %>%
  filter(id %in% soc_layout$id) %>%
  ggraph(layout = "nicely", 
         x = soc, y = soc_layout$y) +
    geom_edge_link(aes(alpha = alpha*0.4, color = week)) +
    scale_edge_color_continuous(low = "blue", high = "dark red") +
    theme_graph() + 
    theme(legend.position = "none")
```

``` {r draw_site_structure}
soc_cc_plot + 
  ggtitle('Social Psychology Site Structure')
bio_cc_plot +
  ggtitle('Subject A Site Structure')
```

# Just add humans

## Cluster Summaries

``` {r cluster_summaries}
cluster_summary_table <- function(x) {
  x %>%
    filter(!is.na(cluster)) %>% 
    summarise(n = n(), 
              Grade = mean(Grade, na.rm = T),
              Accesses = mean(Accesses, na.rm = T),
              clicks_pa = mean(mean_clicks_per_access, na.rm = T),
              sd_clicks = mean(sd_clicks, na.rm = T),
              median_time_pa = as.duration(mean(median_time_per_access)),
              sd_time = as.duration(mean(sd_time)),
              TotalTime = as.duration(mean(TotalTime)),
              ForumViews = mean(ForumViews)) %>% 
    flextable() %>% 
    autofit()
}


cluster_plot <- function(x) {
  p <- x %>% 
    filter(Role == "S") %>% 
    ggplot(aes(x = Dim1, y = Dim2, 
               color = factor(cluster), 
               size = grade_quartile)) +
    scale_size_continuous(trans = "exp",
                          name = "Grade Quartile",
                          guide = "none") +
    geom_point(alpha = 0.4) +
    scale_color_manual(values = wesanderson::wes_palettes[[8]],
                       name = "Cluster") +
    theme_void() 
  p
}

cluster_plot(nodesBio) + ggtitle("Biopsychology Clusters")

nodesBio %>% group_by(cluster) %>% cluster_summary_table()

cluster_plot(nodesSoc) + ggtitle("Social Psychology Clusters")

nodesSoc %>% group_by(cluster) %>% cluster_summary_table()
```

## Site Activity

``` {r draw_site_activity, eval = F}
soc_cluster_act_plot + ggtitle(
  'Social Psychology Site Activity', 
  subtitle = 'By cluster, time running left to right')
bio_cluster_act_plot + ggtitle(
  'Biopsychology Site Activity', 
  subtitle = 'By cluster, time running left to right')
```

## Spectral Analysis

``` {r spectral_analysis_functions}
# spectral analysis
timelist_to_difference_n <- function(timestamps, n = 100) {
  n <- length(timestamps)
  M <- matrix(rep(as.numeric(timestamps), n), ncol = n, nrow = n)
  M_diff <- M - t(M)
  diff <- abs(M_diff[upper.tri(M_diff, diag = FALSE)])
  if (min(diff) == 0) {
    next_min <- min(diff[diff > 0])
    diff <- c(diff[diff > 0], rep(next_min * 0.5, length(diff[diff == 0])))
  }
  output <- lubridate::as.duration(diff)
  if (length(output) > n) {
    output <- sample(output, n)
  }
  return(output)
}

aa_to_session_intervals_n <- function(x, n = 100) {
  x %>% 
    filter(!is.na(cluster)) %>% 
    select(cluster, grade_quartile, person, timestamp, session) %>% 
    group_by(cluster, grade_quartile, person, session) %>% 
    nest() %>% 
    mutate(intervals = map2(.x = data, .y = n, 
                            ~timelist_to_difference_n(.x$timestamp, .y))) %>% 
    select(cluster, grade_quartile, person, session, intervals) %>% 
    unnest() %>% 
    filter(as.duration(intervals) < as.duration('1 day'))
}

aa_to_item_intervals_n <- function(x, n = 100) {
  x %>% 
    filter(!is.na(cluster)) %>% 
    select(cluster, person, timestamp, session, type) %>% 
    group_by(cluster, person, session, type) %>% 
    nest() %>% 
    mutate(intervals = map2(.x = data, .y = n, 
                            ~timelist_to_difference_n(.x$timestamp, .y))) %>% 
    select(cluster, person, session, intervals, type) %>% 
    unnest() %>% 
    filter(as.duration(intervals) < as.duration('1 day'))
}
```

``` {r spectral_analysis_bio_write, eval = FALSE}
set.seed(31415)
bio_act_session_intervals_n <- aa_to_session_intervals_n(aabio)
saveRDS(bio_act_session_intervals_n, 'bio_act_session_intervals_n.rds')
#bio_act_item_intervals_n <- aa_to_item_intervals_n(aabio)
#saveRDS(bio_act_item_intervals_n, 'bio_act_item_intervals_n.rds')
```

``` {r spectral_analysis_bio_read}
bio_act_session_intervals_n <- readRDS('bio_act_session_intervals_n.rds') 

bio_cluster_comparison_session <- bio_act_session_intervals_n %>% 
  plot_timestamp_spectrum() +
  facet_grid(cluster ~ .)+
  ggtitle('Biopsychology Activity Spectrums', subtitle = 'Signals below 1 day only')

bio_cluster_comparison_session_grades <- bio_act_session_intervals_n %>% 
  plot_timestamp_spectrum(color = cluster) +
  facet_grid(grade_quartile ~ .)+
  ggtitle('Biopsychology Activity Spectrums', subtitle = 'Signals below 1 day only')

# bio_act_item_intervals_n <- readRDS('bio_act_item_intervals_n.rds')
# bio_act_item_plot <- bio_act_item_intervals %>% 
  #filter(type != "db_collection_entry") %>% 
  # plot_timestamp_spectrum() +
  # theme(strip.text.y = element_text(angle = 0))+
  # facet_grid(type ~ cluster) + 
  # ggtitle('Activity spectrums - Biopsychology', subtitle = 'Cluster v Item')
```

``` {r spectral_analysis_soc_write, eval = FALSE}

soc_act_session_intervals_n <- aa_to_session_intervals_n(aasoc)
saveRDS(soc_act_session_intervals_n, 'soc_act_session_intervals_n.rds')
```

``` {r spectral_analysis_soc_read}
soc_act_session_intervals_n <- readRDS('soc_act_session_intervals_n.rds')

soc_cluster_comparison_session <- soc_act_session_intervals_n %>% 
  plot_timestamp_spectrum() +
  facet_grid(grade_quartile ~ .) +
  ggtitle('Social Psychology Activity Spectrums', subtitle = 'Signals above 1 day removed')

```

``` {r draw_spetral_plots}
#soc_cluster_spectral_comparison
soc_cluster_comparison_session
#bio_cluster_spectral_comparison
bio_cluster_comparison_session
```

``` {r item_spectrums}
# In progress
```