---
title: "Analysis of Engagement in Social Psychology and Biopsychology Courses - Qual Participants"
author: "Ben Hicks"
output: html_document
---

The following data is from the Social Psychology (PSY203 and PSY453) and Biopsychology (PSY208 and PSY408) subjects run in 2018 Session 2. The aim of this report is to look at student engagement in these courses through the data available in Blackboard Learn. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
#knitr::opts_chunk$set(dev = 'png') - - uncomment to reduce file size
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(GGally))
suppressPackageStartupMessages(library(network))
suppressPackageStartupMessages(library(tidygraph))
suppressPackageStartupMessages(library(ggraph))
suppressPackageStartupMessages(library(ggthemes))
suppressPackageStartupMessages(library(lakit))
suppressPackageStartupMessages(library(dendextend))
suppressPackageStartupMessages(library(cluster))
suppressPackageStartupMessages(library(factoextra))
suppressPackageStartupMessages(library(ggsci))
```

``` {r importing, cache=TRUE}
source(file.path('..', 'R', 'psych_functions.R'))
source(file.path('..', 'R', 'import-data.R'))
```


# Qual Participant Summaries

``` {r qual_summary}
# feed in grouped_by df
cluster_summary_table <- function(x) {
  x %>%
    summarise(n = n(), 
              Biopsych = sum(Subject == "Biopsychology"),
              SocPsych = sum(Subject == "Social Psychology"),
              Grade = mean(Grade, na.rm = T),
              Grade_sd = sd(Grade, na.rm = T),
              HDs = sum(Grade > 80),
              Accesses = mean(Accesses),
              clicks_pa = mean(mean_clicks_per_access),
              sd_clicks = mean(sd_clicks),
              median_time_pa = as.duration(mean(median_time_per_access)),
              sd_time = as.duration(mean(sd_time)),
              TotalTime = as.duration(mean(TotalTime)),
              ForumViews = mean(ForumViews)) %>% 
    knitr::kable()
}

nodes %>% group_by(qual_participant) %>% cluster_summary_table()

```


# Activity

## Item Activity

We can examine student interactions with the course content at a finer level, and filter according to activities of note.

### Social Psychology

``` {r social_psych_usage, cache = TRUE}
# Using the ACTIVITY_ACCUMULATOR data to examine item activity
qualNames <- unique(nodes[nodes$qual_participant,]$Name)
socUsage <- aaSoc18 %>%
  select(name, timestamp, content, role, forum_bool) %>%
  filter(role == "S" & timestamp < as.Date('2018-11-01')) %>%
  filter(name %in% qualNames) %>% 
  mutate(Activity = if_else(
    str_detect(content, "Week 2: The Social Self|Week 3:Perc|Week 4:Stereo|Week 5:Att|Week 6:Conf|Week 10:Att|Week 11:Help|Applied Social"), 
    "Tutorial",
    if_else(forum_bool, "Forum", "Other")
    )
    )%>%
  mutate(Activity = if_else(is.na(Activity), "Other", Activity)) %>%
  rename(Name = name) %>% 
  left_join(nodesSoc %>% 
          select(Name, Accesses, TotalTime, grade_quartile, ForumViews) %>% 
          unique(), by = c("Name"))

socGameActPlot <- socUsage %>% 
  ggplot(aes(x = timestamp, y = grade_quartile, color = ForumViews)) +
  scale_color_gradient2(name = "Forum Participation", 
                         low = "Black",
                          mid = "Dark Blue",
                         high = "Red",
                         midpoint = 3.5,
                         trans = "log1p",
                         breaks = c(2, 10, 50, 250, 1250)) +  
  geom_jitter(alpha = 0.05) +
  scale_y_continuous(name = "Grade Quartile") +
  ggthemes::theme_tufte() + 
  theme(axis.title.x = element_blank()) + 
  facet_grid(Activity ~ .) +
  ggtitle("Usage of items in Social Psychology - Qual Participants")

socGameActPlot

```

``` {r soc_site_nav, cache = T}
socSiteGraphList <- build_site_activity_graph_list(
  cc_nodes = nodes_from_cc(ccSoc18, subject = "Soc"),
  cc_edges = edges_from_cc(ccSoc18),
  aa_nodes = nodes_from_aa(aaSoc18),
  aa_edges = edges_from_aa(aaSoc18))

draw_site_graph(socSiteGraphList) + ggtitle("Social Psychology site structure")

socQualGraphList <- build_site_activity_graph_list(
  cc_nodes = nodes_from_cc(ccSoc18, subject = "Soc"),
  cc_edges = edges_from_cc(ccSoc18),
  aa_nodes = nodes_from_aa(aaSoc18 %>% filter(qual_participant)),
  aa_edges = edges_from_aa(aaSoc18 %>% filter(qual_participant)))

draw_site_activity_graph(socQualGraphList, alpha_low = 0.05) + ggtitle("Social Psychology activity of qual students")

```

\newpage

### Biopsychology Activity

``` {r biopsych_usage, cache = TRUE}
# Using the ACTIVITY_ACCUMULATOR data to examine item activity

bioUsage <- aaBio18 %>%
  select(name, timestamp, content, role, forum_bool) %>%
  filter(role == "S" , name %in% qualNames) %>%
  #& timestamp < as.Date('2018-11-01')) %>%
  mutate(Activity = if_else(str_detect(tolower(content), "game"), "Game",
                                   if_else(str_detect(tolower(content), "in the lab"), "In the lab",
                                           if_else(forum_bool, "Forum", "Other")))) %>%
  mutate(Activity = if_else(is.na(Activity), "Other", Activity)) %>%
  rename(Name = name) %>%
  merge(nodesBio %>% select(Name, Accesses, grade_quartile, ForumViews, 
                            mean_clicks_per_access, sd_clicks, median_time_per_access, sd_time,
                            TotalTime), all.x = T)

bioGameActPlotFocus <- bioUsage %>% filter(timestamp < as.Date('2018-10-15')) %>%
  ggplot(aes(x = timestamp, y = grade_quartile, color = ForumViews)) +
  scale_color_gradient2(name = "Forum Participation", 
                         low = "Black",
                          mid = "Dark Blue",
                         high = "Red",
                         midpoint = 3.5,
                         trans = "log1p",
                         breaks = c(2, 10, 50, 250, 1250)) +  
  geom_jitter(alpha = 0.1) +
  scale_y_continuous(name = "Grade Quartile") +
  ggthemes::theme_tufte() + 
  theme(axis.title.x = element_blank()) + 
  facet_grid(Activity ~ .) +
  ggtitle("Usage of items in Biopsychology - Qual Participants")

bioGameActPlotFocus
```


``` {r bio_site_nav, cache = T}
bioSiteGraphList <- build_site_activity_graph_list(
  cc_nodes = nodes_from_cc(ccBio18, subject = "Bio"),
  cc_edges = edges_from_cc(ccBio18),
  aa_nodes = nodes_from_aa(aaBio18),
  aa_edges = edges_from_aa(aaBio18))

draw_site_graph(bioSiteGraphList) + ggtitle("Biopsychology site structure")

bioQualGraphList <- build_site_activity_graph_list(
  cc_nodes = nodes_from_cc(ccBio18, subject = "Bio"),
  cc_edges = edges_from_cc(ccBio18),
  aa_nodes = nodes_from_aa(aaBio18 %>% filter(qual_participant)),
  aa_edges = edges_from_aa(aaBio18 %>% filter(qual_participant)))

draw_site_activity_graph(bioQualGraphList, alpha_low = 0.05) + ggtitle("Biopsychology activity of qual students")
```




