---
title: "Analysis of Engagement in Social Psychology and Biopsychology Courses - Clustering"
author: "Ben Hicks"
output: html_document
---

The following data is from the Social Psychology (PSY203 and PSY453) and Biopsychology (PSY208 and PSY408) subjects run in 2018 Session 2. The aim of this report is to look at student engagement in these courses through the data available in Blackboard Learn. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
#knitr::opts_chunk$set(dev = 'png') - - uncomment to reduce file size
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(GGally))
suppressPackageStartupMessages(library(network))
suppressPackageStartupMessages(library(tidygraph))
suppressPackageStartupMessages(library(ggraph))
suppressPackageStartupMessages(library(ggthemes))
suppressPackageStartupMessages(library(lakit))
suppressPackageStartupMessages(library(dendextend))
suppressPackageStartupMessages(library(cluster))
suppressPackageStartupMessages(library(factoextra))
suppressPackageStartupMessages(library(ggsci))
```

``` {r importing, cache=TRUE}
source(file.path('..','R','psych_functions.R'))
DATDIR <- file.path("~", "Data", "Psychology")
BIOAAFILE <- "aa Biopsychology 201860.csv"
SOCAAFILE <- "aa Social Psychology 201860.csv"
MARKSFILE <- "marks Psychology 201860.csv"
BIOCCFILE <- "cc Biopsychology 201860.csv"
SOCCCFILE <- "cc Social Psychology 201860.csv"
POSTSFILE <- "posts Psychology 201860.csv"
QUALFILE <- "qual participants.csv"
ZERODATE <- "2018-05-01" # pre-session date used to colour the site activity structure paths. Affects animation time before student activity

snData <- read_csv(file.path(DATDIR, POSTSFILE), col_types = "cccffn")
# Activity Accumulator Data --------------------------------------
tidy_aa <- function(df) {
  df <- df %>%
    filter(!is.na(id), role == "S") %>%
    mutate(forum_bool = str_detect(tolower(data), "discus|thread|forum"),
                                      content = tidyDataField(data),
                                      name = paste0(lastname, ', ', firstname)) %>%
    rename(session = session_id, id = s_id, login_id = id)
  df <- df %>% mutate(forum_bool = if_else(is.na(forum_bool), FALSE, forum_bool))
  return(df)
} # --------------------------------------------------------------
aaSoc18 <- read_csv(file.path(DATDIR, SOCAAFILE), col_types = "ffTcccccffcccf") # can cause performance hit, if so do not specity columns and add as.factor lines to tidy_aa 
aaSoc18 <- tidy_aa(aaSoc18)

aaBio18 <- read_csv(file.path(DATDIR, BIOAAFILE), col_types = "ffTcccccffcccf")
aaBio18 <- tidy_aa(aaBio18)

# Course contents
tidy_cc <- function(x) {
  x %>% 
    select(pk1, parent_pk1, title, data = main_data) %>% 
    mutate(pk1 = factor(pk1), parent_pk1 = factor(parent_pk1)) %>% 
    mutate(title = str_trunc(title, width = 45))
}
ccSoc18 <- read_csv(file.path(DATDIR, SOCCCFILE)) %>% tidy_cc()
ccBio18 <- read_csv(file.path(DATDIR, BIOCCFILE)) %>% tidy_cc()


# Node preparation (i.e. data for each student)

# Mark data
marks18 <- read_csv(file.path(DATDIR, MARKSFILE)) %>%
  mutate(mark = score / possible, id = factor(id)) 
marksSummary18 <- marks18 %>% 
  group_by(subject_name, student, id) %>%
  summarise(avg_grade = mean(mark, na.rm = T)) %>%
  rename(name = student, subject = subject_name)
marksSoc18 <- marksSummary18 %>% filter(subject == "Social Psychology")
marksBio18 <- marksSummary18 %>% filter(subject == "Biopsychology")

# Discussion views from aa -------------------------------------------
discussion_views <- function(df) {
  df %>%
    select(name, subject, forum_bool, data) %>%
    filter(  forum_bool |
           str_detect(tolower(data), "discus|thread|forum")
         ) %>% 
    group_by(subject, name) %>%
    summarise(ForumViews = n()) %>%
    group_by(name) %>%
    summarise(ForumViews = sum(ForumViews))
} # ------------------------------------------------------------------
discussionViewsBio <- discussion_views(aaBio18)
discussionViewsSoc <- discussion_views(aaSoc18)

# Other interaction data from aa
aaSummaryBio <- aa_summary_by_user(aaBio18)
aaSummaryBio <- merge(aaSummaryBio, unique(aaBio18 %>% select(id, name)))
aaSummarySoc <- aa_summary_by_user(aaSoc18)
aaSummarySoc <- merge(aaSummarySoc, unique(aaSoc18 %>% select(id, name)))


# Node assembly ----------------------------------------------------------
node_merge_and_make <- function(aaData, marksData, discussionsData) {
  df <- aaData %>% select(-id) %>%
  merge(marksData) %>%
  merge(discussionsData, all.x = T) %>%
  select(ID = id, Name = name, Subject = subject, 
         Grade = avg_grade, Accesses = accesses, 
         mean_clicks_per_access, sd_clicks, 
         median_time_per_access, sd_time, 
         TotalTime = total_time,
         ForumViews) %>%
    mutate(grade_quartile = ntile(Grade, 4), ID = factor(ID))
  if (any(is.na(df$ForumViews))) {
    df[is.na(df$ForumViews),]$ForumViews <- 0
  }
  if (any(is.na(df$Accesses))) {
    df[is.na(df$Accesses),]$Accesses <- 0  
  }
  if (any(is.na(df$TotalTime))) {
    df[is.na(df$TotalTime),]$TotalTime <- 0
  }
  df$Role <- "S"
  return(df)
} # ----------------------------------------------------------------------
nodesBio <- node_merge_and_make(aaSummaryBio, marksBio18, discussionViewsBio)
nodesSoc <- node_merge_and_make(aaSummarySoc, marksSoc18, discussionViewsSoc)

# -----------------------------------------------------------------------------
subject_code_to_name <- function(x) {
  ifelse(str_detect(x, "208|458"), "Biopsychology", "Social Psychology")
} # ---------------------------------------------------------------------------

## ADD ##
allStudentIDs <- unique(marks18$id)
allForumIDs <- unique(snData$msgfrom)
teacherIDs <- setdiff(allForumIDs, allStudentIDs)
teacherNodes <- snData %>% 
  filter(msgfrom %in% teacherIDs) %>% 
  select(Name = poster, ID = msgfrom, Subject = subject) %>% 
  mutate(ID = as.factor(ID), 
         Grade = 0, Accesses = 100, ForumViews = 200, Role = "T", Subject = subject_code_to_name(Subject), 
         TotalTime = as.duration(200000), median_time_per_access = as.duration(2000), sd_time = as.duration(200),
         mean_clicks_per_access = 0, sd_clicks = 0, grade_quartile = 0) %>%
  unique()
teacherNodesBio <- teacherNodes %>% filter(Subject == "Biopsychology")
teacherNodesSoc <- teacherNodes %>% filter(Subject == "Social Psychology")

nodesBio <- rbind(nodesBio, teacherNodesBio)
nodesSoc <- rbind(nodesSoc, teacherNodesSoc)

# Edge assembly
edgesAll <- data.frame(from = snData$msgfrom, 
                       to = snData$msgto, 
                       weight = snData$strength, 
                       Subject = snData$subject) %>%
  mutate(Subject = subject_code_to_name(Subject))
questions <- edgesAll %>% 
  filter(to == "NULL") %>% 
  group_by(from, Subject) %>% 
  summarise(Questions = sum(weight))
edges <- edgesAll %>% filter(to != "NULL")
edgesBio <- edges %>% filter(str_detect(Subject, "Bio"))
edgesSoc <- edges %>% filter(str_detect(Subject, "Social"))


# Cluster analysis
make_normed_matrix <- function(df, return_ids = FALSE) {
  ids <- df$ID
  df$median_time_per_access <- as.numeric(as_duration_from_string(df$median_time_per_access))
  df$sd_time <- as.numeric(as_duration_from_string(df$sd_time))
  df <- as.data.frame(sapply(df %>% 
                         filter(Role == "S") %>% 
                         select(-ID, -Name, -Subject, -Role, -grade_quartile), 
                       as.numeric))
  #gq <- df$grade_quartile # does not need to be normed
  df <- scale(df)
  if (return_ids) {
    df$ID <- ids
  }
  #df$grade_quartile <- gq
  return(df)
}

nodeMatrixBio <- make_normed_matrix(nodesBio)
nodeMatrixSoc <- make_normed_matrix(nodesSoc)
nodes <- rbind(nodesBio, nodesSoc) %>% filter(Role == "S")
nodeMatrix <- make_normed_matrix(nodes)
dAct <- nodeMatrix[, 2:8]

kmed <- pam(dAct, 5)

nodes <- bind_cols(nodes, data.frame(cluster = kmed$cluster))

# Adding Qual participants
qual <- read_csv(file.path(DATDIR, QUALFILE))
qual <- qual %>% 
  rename(qual = Method) %>% 
  select(-Subject)
nodes <- nodes %>% left_join(qual, by = c("Name"))
nodes$qual_participant <- !is.na(nodes$qual)
nodes[is.na(nodes$qual), ]$qual <- "None"

# Joining cluster and qual back with subject ndoes
nodesBio <- left_join(nodesBio, nodes %>% select(ID, Name, cluster, qual))
nodesSoc <- left_join(nodesSoc, nodes %>% select(ID, Name, cluster, qual))

# Joining grade_quartile, cluster and qual back with aa 
nodeInfo <- nodes %>% 
  select(name = Name, grade_quartile, cluster, qual, qual_participant)
aaBio18 <- left_join(aaBio18, nodeInfo)
aaSoc18 <- left_join(aaSoc18, nodeInfo)

# Making graph structures
networkBio <- tbl_graph(nodes = nodesBio, edges = edgesBio) 
networkSoc <- tbl_graph(nodes = nodesSoc, edges = edgesSoc)

```

``` {r site_nav_functions, eval = T}
# Looking at how students navigated the course
edges_from_aa <- function(x) {
  aa <- x %>% 
    filter(role == 'S', !is.na(content_pk1), timestamp < as.POSIXct('2018-10-30')) %>% 
    select(id, time = timestamp, content_pk1, session, 
           grade_quartile, cluster, qual) %>% 
    mutate(id = factor(id), content_pk1 = factor(content_pk1)) %>% 
    arrange(id, time)
  aa_edges <- tibble(from = aa$content_pk1[1:(nrow(aa) - 1)],
                   to =   aa$content_pk1[2:nrow(aa)],
                   time = aa$time[2:nrow(aa)],
                   student_from = aa$id[1:(nrow(aa) - 1)],
                   student = aa$id[2:nrow(aa)],
                   session_from = aa$session[1:(nrow(aa) - 1)],
                   session = aa$session[2:nrow(aa)],
                   grade_quartile = aa$grade_quartile[1:(nrow(aa) - 1)],
                   cluster = aa$cluster[1:(nrow(aa) - 1)],
                   qual = aa$qual[1:(nrow(aa) - 1)])
  aa_edges <- aa_edges %>% 
    filter(student == student_from, session == session_from) %>% 
    select(from, to, time, student, grade_quartile, cluster, qual) %>% 
    mutate(weight = 2)
  return(aa_edges)
}

edges_from_cc <- function(x) {
  cc_edges <- x %>% 
    select(from = parent_pk1, to = pk1) %>% 
    filter(!is.na(from)) %>% 
    mutate(weight = 1, time = as.POSIXct(ZERODATE)) # used manual set time - needs to be well before other data
  return(cc_edges)
}

nodes_from_cc <- function(x, subject) {
  x <- x %>% select(id = pk1, title, data) %>% 
    mutate(forum_bool = str_detect(tolower(data), "discus|thread|forum")) %>% 
    mutate(forum_bool = if_else(is.na(forum_bool), FALSE, forum_bool))
  if (str_detect(subject, "Bio|bio")) {
    x <- x %>% 
      mutate(PageType = if_else(str_detect(title, "lab|Lab"), "Lab", 
                                if_else(forum_bool, "Forum", "Other"))) %>%
      mutate(PageType = if_else(is.na(PageType), "Other", PageType)) %>% 
      mutate(PageType = factor(PageType, levels = c("Forum", "Lab", "Other")))
  }
  if (str_detect(subject, "Soc|soc")) {
   x <- x %>% 
     mutate(PageType = if_else(str_detect(title, "Week 2: The Social Self|Week 3:Perc|Week 4:Stereo|Week 5:Att|Week 6:Conf|Week 10:Att|Week 11:Help|Applied Social"), 
                               "Tutorial",
                               if_else(forum_bool, "Forum", "Other"))) %>%
     mutate(PageType = if_else(is.na(PageType), "Other", PageType)) %>% 
     mutate(PageType = factor(PageType, levels = c("Forum", "Tutorial", "Other")))
   }
  return(x %>% select(-data))
  }

nodes_from_aa <- function(x) {
  x %>% 
    filter(!is.na(content_pk1)) %>% 
    left_join(nodes %>% select(id = ID, ))
    select(id = content_pk1, cluster, qual) %>% # might need to calc grade quartile
    unique()
}

build_site_activity_graph_list <- function(cc_nodes, cc_edges, aa_nodes, aa_edges) {
  cc_graph <- tbl_graph(nodes = cc_nodes, edges = cc_edges)
  cc_layout <- create_layout(cc_graph, layout = "nicely")
  np <- data.frame(x = cc_layout$x, y = cc_layout$y)
  aa_graph <- tbl_graph(nodes = cc_nodes, edges = aa_edges)
  joined_graph <- cc_graph %>% graph_join(aa_graph) 
  joined_graph <- joined_graph %>% 
    activate(edges) %>% 
    mutate(student = if_else(is.na(student), "_course structure", as.character(student)))
  return(list(joined_graph, np))
}


draw_site_graph <- function(graph_list){
  graph <- graph_list[[1]] %>% 
    activate(edges) %>% 
    filter(weight == 1)
  np <- graph_list[[2]]
  g <- ggraph(graph, layout = "manual", node.positions = np) +
    geom_node_point(aes(color = PageType)) + 
    geom_node_text(aes(label = title), alpha = 0.4, repel = T, size = 1) +
    geom_edge_link0() +
    theme_graph() +
    theme(#legend.position = "none",
        text = element_text(family = "TT Times New Roman"))
  return(g)
}

draw_site_activity_graph <- function(graph_list, alpha_low = 0.1) {
  g <- ggraph(graph_list[[1]], layout = "manual", node.positions = graph_list[[2]]) +
    geom_node_point() + 
    geom_edge_link0(aes(color = as.numeric(time), 
                      width = weight, 
                      alpha = 2 - weight,
                      group = time)) +
    scale_edge_alpha_continuous(range = c(alpha_low, 1)) +
    scale_edge_width(range = c(0.6,1)) +
    scale_edge_color_gradientn(colors = c("black", "#47ADED", "#D81C41")) +
    theme_graph() +
    theme(legend.position = "none",
        text = element_text(family = "TT Times New Roman"))
  return(g)
}
```



# Modelling

## Cluster Analysis

Clustering seemed to be most parsimonious choosing the *k-medoids* algorithm with 5 clusters across both subjects, and *excluding academic data*. See the document *cluster analysis psych* for more detail on why this was chosen.

``` {r modelling_setup}
# feed in grouped_by df
cluster_summary_table <- function(x) {
  x %>%
    summarise(n = n(), 
              Biopsych = sum(Subject == "Biopsychology"),
              SocPsych = sum(Subject == "Social Psychology"),
              Grade = mean(Grade, na.rm = T),
              Grade_sd = sd(Grade, na.rm = T),
              HDs = sum(Grade > 80, na.rm = T),
              Accesses = mean(Accesses, na.rm = T),
              clicks_pa = mean(mean_clicks_per_access, na.rm = T),
              sd_clicks = mean(sd_clicks, na.rm = T),
              median_time_pa = as.duration(mean(median_time_per_access)),
              sd_time = as.duration(mean(sd_time)),
              TotalTime = as.duration(mean(TotalTime)),
              ForumViews = mean(ForumViews)) %>% 
    knitr::kable()
}

fviz_cluster(kmed)

nodes %>% group_by(cluster) %>% cluster_summary_table()

```

## Activity by Cluster

### Social Psychology

``` {r social_psych_usage, cache = TRUE}
# Using the ACTIVITY_ACCUMULATOR data to examine item activity
socUsage <- aaSoc18 %>%
  select(name, timestamp, content, role, forum_bool) %>%
  filter(role == "S" & timestamp < as.Date('2018-11-01')) %>%
  mutate(Activity = if_else(
    str_detect(content, "Week 2: The Social Self|Week 3:Perc|Week 4:Stereo|Week 5:Att|Week 6:Conf|Week 10:Att|Week 11:Help|Applied Social"), 
    "Tutorial",
    if_else(forum_bool, "Forum", "Other")
    )
    )%>%
  mutate(Activity = if_else(is.na(Activity), "Other", Activity)) %>%
  rename(Name = name) %>% 
  left_join(nodes %>% filter(Subject == "Social Psychology") %>% 
          select(Name, Accesses, TotalTime, grade_quartile, cluster) %>% 
          unique(), by = c("Name")) %>% 
  mutate(cluster = factor(cluster))

socGameActPlot <- socUsage %>% 
  ggplot(aes(x = timestamp, y = grade_quartile, color = cluster)) +
  geom_jitter(alpha = 0.015) +
  scale_y_continuous(name = "Grade Quartile") +
  ggthemes::theme_tufte() + 
  theme(axis.title.x = element_blank()) + 
  facet_grid(Activity ~ .) +
  ggtitle("Usage of items in Social Psychology")

socGameActPlot
```




``` {r soc_site_nav, cache = T}
socSiteGraphList <- build_site_activity_graph_list(
  cc_nodes = nodes_from_cc(ccSoc18, subject = "Soc"),
  cc_edges = edges_from_cc(ccSoc18),
  aa_nodes = nodes_from_aa(aaSoc18),
  aa_edges = edges_from_aa(aaSoc18))

draw_site_graph(socSiteGraphList) + ggtitle("Social Psychology site structure")

draw_site_activity_graph(socSiteGraphList, alpha_low = 0.02) + facet_wrap(~ cluster) + ggtitle("Social Psychology activity by cluster")
```

\newpage

### Biopsychology

``` {r biopsych_usage, cache = TRUE}
# Using the ACTIVITY_ACCUMULATOR data to examine item activity

bioUsage <- aaBio18 %>%
  select(name, timestamp, content, role, forum_bool) %>%
  filter(role == "S" ) %>%
  #& timestamp < as.Date('2018-11-01')) %>%
  mutate(Activity = if_else(str_detect(tolower(content), "game"), "Game",
                                   if_else(str_detect(tolower(content), "in the lab"), "In the lab",
                                           if_else(forum_bool, "Forum", "Other")))) %>%
  mutate(Activity = if_else(is.na(Activity), "Other", Activity)) %>%
  rename(Name = name) %>%
  merge(nodes %>% filter(Subject == "Biopsychology") %>%  
          select(Name, Accesses, grade_quartile, cluster, 
                            mean_clicks_per_access, sd_clicks, median_time_per_access, sd_time,
                            TotalTime), all.x = T) %>% 
  mutate(cluster = factor(cluster))

bioGameActPlotFocus <- bioUsage %>% filter(timestamp < as.Date('2018-10-15')) %>%
  ggplot(aes(x = timestamp, y = grade_quartile, color = cluster)) +
  geom_jitter(alpha = 0.01) +
  scale_y_continuous(name = "Grade Quartile") +
  ggthemes::theme_tufte() + 
  theme(axis.title.x = element_blank()) + 
  facet_grid(Activity ~ .) +
  ggtitle("Usage of items in Biopsychology with clusters")

bioGameActPlotFocus
```

``` {r bio_site_nav, cache = T}
bioSiteGraphList <- build_site_activity_graph_list(
  cc_nodes = nodes_from_cc(ccBio18, subject = "Bio"),
  cc_edges = edges_from_cc(ccBio18),
  aa_nodes = nodes_from_aa(aaBio18),
  aa_edges = edges_from_aa(aaBio18))

draw_site_graph(bioSiteGraphList) + ggtitle("Biopsychology site structure")

draw_site_activity_graph(bioSiteGraphList, alpha_low = 0.02) + facet_wrap(~ cluster) + ggtitle("Biopsychology activity by cluster")
```

\newpage

## Activity profile of clusters

In progress - spectral analysis of clusters

``` {r spectral_analysis_of_timestamps, eval = F}
bioStampsAndClusters <- inner_join(nodes%>% 
                                     select(name = Name, cluster),
                                aaBio18 %>% 
                                  select(name, timestamp, session))

c1sample <- sample(bioStampsAndClusters[bioStampsAndClusters$cluster == 1,]$name, 5)
c2sample <- sample(bioStampsAndClusters[bioStampsAndClusters$cluster == 2,]$name, 5)
c3sample <- sample(bioStampsAndClusters[bioStampsAndClusters$cluster == 3,]$name, 5)
c4sample <- sample(bioStampsAndClusters[bioStampsAndClusters$cluster == 4,]$name, 5)
c5sample <- sample(bioStampsAndClusters[bioStampsAndClusters$cluster == 5,]$name, 5)

map(c1sample, function(x)timelist_to_difference(bioStampsAndClusters %>% filter(name == x)))

c11 <- data.frame(timelist_to_difference(bioStampsAndClusters[bioStampsAndClusters$name == c1sample[[1]], ]))
plot_timestamp_spectrum(tl, group = id) + facet_wrap(~ id)
```

In progress - animated activity by cluster

``` {r animated_activity_of_clusters, eval = F}
```

\newpage

## Network analysis on forum posts and replies, highlighting clusters

### Social Psychology

``` {r fig.height = 4}
# Graphs for Social Psychology 
gSocCluster <- buildClusterKMetGraph(networkSoc %>% 
                                       activate(nodes) %>% 
                                       filter(Role == "S"))
gSocCluster + ggtitle("Clusters for PSY203: Social Psychology")
```

\newpage

### Biopsychology

``` {r fig.height = 4}
gBioCluster <- buildClusterKMetGraph(networkBio %>% 
                                       activate(nodes) %>% 
                                       filter(Role == "S"))
gBioCluster + ggtitle("Clusters for PSY208: Biopsychology")
```
