---
title: "Analysis of Engagement in Social Psychology and Biopsychology Courses - Forum Activity"
author: "Ben Hicks"
output: html_document
---

The following data is from the Social Psychology (PSY203 and PSY453) and Biopsychology (PSY208 and PSY408) subjects run in 2018 Session 2. The aim of this report is to look at student engagement in these courses through the data available in Blackboard Learn. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F, cache = T)
#knitr::opts_chunk$set(dev = 'png') - - uncomment to reduce file size
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(GGally))
suppressPackageStartupMessages(library(network))
suppressPackageStartupMessages(library(tidygraph))
suppressPackageStartupMessages(library(ggraph))
suppressPackageStartupMessages(library(ggthemes))
suppressPackageStartupMessages(library(lakit))
suppressPackageStartupMessages(library(dendextend))
suppressPackageStartupMessages(library(cluster))
suppressPackageStartupMessages(library(factoextra))
suppressPackageStartupMessages(library(ggsci))
```

``` {r importing, cache=TRUE}
source(file.path('..','R','psych_functions.R'))
DATDIR <- file.path("~", "Data", "Psychology")
BIOAAFILE <- "aa Biopsychology 201860.csv"
SOCAAFILE <- "aa Social Psychology 201860.csv"
MARKSFILE <- "marks Psychology 201860.csv"
BIOCCFILE <- "cc Biopsychology 201860.csv"
SOCCCFILE <- "cc Social Psychology 201860.csv"
POSTSFILE <- "posts Psychology 201860.csv"
QUALFILE <- "qual participants.csv"
ZERODATE <- "2018-05-01" # pre-session date used to colour the site activity structure paths. Affects animation time before student activity

snData <- read_csv(file.path(DATDIR, POSTSFILE), col_types = "cccffn")
# Activity Accumulator Data --------------------------------------
tidy_aa <- function(df) {
  df <- df %>%
    filter(!is.na(id), role == "S") %>%
    mutate(forum_bool = str_detect(tolower(data), "discus|thread|forum"),
                                      content = tidyDataField(data),
                                      name = paste0(lastname, ', ', firstname)) %>%
    rename(session = session_id, id = s_id, login_id = id)
  df <- df %>% mutate(forum_bool = if_else(is.na(forum_bool), FALSE, forum_bool))
  return(df)
} # --------------------------------------------------------------
aaSoc18 <- read_csv(file.path(DATDIR, SOCAAFILE), col_types = "ffTcccccffcccf") # can cause performance hit, if so do not specity columns and add as.factor lines to tidy_aa 
aaSoc18 <- tidy_aa(aaSoc18)

aaBio18 <- read_csv(file.path(DATDIR, BIOAAFILE), col_types = "ffTcccccffcccf")
aaBio18 <- tidy_aa(aaBio18)

# Course contents
tidy_cc <- function(x) {
  x %>% 
    select(pk1, parent_pk1, title, data = main_data) %>% 
    mutate(pk1 = factor(pk1), parent_pk1 = factor(parent_pk1)) %>% 
    mutate(title = str_trunc(title, width = 45))
}
ccSoc18 <- read_csv(file.path(DATDIR, SOCCCFILE)) %>% tidy_cc()
ccBio18 <- read_csv(file.path(DATDIR, BIOCCFILE)) %>% tidy_cc()


# Node preparation (i.e. data for each student)

# Mark data
marks18 <- read_csv(file.path(DATDIR, MARKSFILE)) %>%
  mutate(mark = score / possible, id = factor(id)) 
marksSummary18 <- marks18 %>% 
  group_by(subject_name, student, id) %>%
  summarise(avg_grade = mean(mark, na.rm = T)) %>%
  rename(name = student, subject = subject_name)
marksSoc18 <- marksSummary18 %>% filter(subject == "Social Psychology")
marksBio18 <- marksSummary18 %>% filter(subject == "Biopsychology")

# Discussion views from aa -------------------------------------------
discussion_views <- function(df) {
  df %>%
    select(name, subject, forum_bool, data) %>%
    filter(  forum_bool |
           str_detect(tolower(data), "discus|thread|forum")
         ) %>% 
    group_by(subject, name) %>%
    summarise(ForumViews = n()) %>%
    group_by(name) %>%
    summarise(ForumViews = sum(ForumViews))
} # ------------------------------------------------------------------
discussionViewsBio <- discussion_views(aaBio18)
discussionViewsSoc <- discussion_views(aaSoc18)

# Other interaction data from aa
aaSummaryBio <- aa_summary_by_user(aaBio18)
aaSummaryBio <- merge(aaSummaryBio, unique(aaBio18 %>% select(id, name)))
aaSummarySoc <- aa_summary_by_user(aaSoc18)
aaSummarySoc <- merge(aaSummarySoc, unique(aaSoc18 %>% select(id, name)))


# Node assembly ----------------------------------------------------------
node_merge_and_make <- function(aaData, marksData, discussionsData) {
  df <- aaData %>% select(-id) %>%
  merge(marksData) %>%
  merge(discussionsData, all.x = T) %>%
  select(ID = id, Name = name, Subject = subject, 
         Grade = avg_grade, Accesses = accesses, 
         mean_clicks_per_access, sd_clicks, 
         median_time_per_access, sd_time, 
         TotalTime = total_time,
         ForumViews) %>%
    mutate(grade_quartile = ntile(Grade, 4), ID = factor(ID))
  if (any(is.na(df$ForumViews))) {
    df[is.na(df$ForumViews),]$ForumViews <- 0
  }
  if (any(is.na(df$Accesses))) {
    df[is.na(df$Accesses),]$Accesses <- 0  
  }
  if (any(is.na(df$TotalTime))) {
    df[is.na(df$TotalTime),]$TotalTime <- 0
  }
  df$Role <- "S"
  return(df)
} # ----------------------------------------------------------------------
nodesBio <- node_merge_and_make(aaSummaryBio, marksBio18, discussionViewsBio)
nodesSoc <- node_merge_and_make(aaSummarySoc, marksSoc18, discussionViewsSoc)

# -----------------------------------------------------------------------------
subject_code_to_name <- function(x) {
  ifelse(str_detect(x, "208|458"), "Biopsychology", "Social Psychology")
} # ---------------------------------------------------------------------------

## ADD ##
allStudentIDs <- unique(marks18$id)
allForumIDs <- unique(snData$msgfrom)
teacherIDs <- setdiff(allForumIDs, allStudentIDs)
teacherNodes <- snData %>% 
  filter(msgfrom %in% teacherIDs) %>% 
  select(Name = poster, ID = msgfrom, Subject = subject) %>% 
  mutate(ID = as.factor(ID), 
         Grade = 0, Accesses = 100, ForumViews = 200, Role = "T", Subject = subject_code_to_name(Subject), 
         TotalTime = as.duration(200000), median_time_per_access = as.duration(2000), sd_time = as.duration(200),
         mean_clicks_per_access = 0, sd_clicks = 0, grade_quartile = 0) %>%
  unique()
teacherNodesBio <- teacherNodes %>% filter(Subject == "Biopsychology")
teacherNodesSoc <- teacherNodes %>% filter(Subject == "Social Psychology")

nodesBio <- rbind(nodesBio, teacherNodesBio)
nodesSoc <- rbind(nodesSoc, teacherNodesSoc)

# Edge assembly
edgesAll <- data.frame(from = snData$msgfrom, 
                       to = snData$msgto, 
                       weight = snData$strength, 
                       Subject = snData$subject) %>%
  mutate(Subject = subject_code_to_name(Subject))
questions <- edgesAll %>% 
  filter(to == "NULL") %>% 
  group_by(from, Subject) %>% 
  summarise(Questions = sum(weight))
edges <- edgesAll %>% filter(to != "NULL")
edgesBio <- edges %>% filter(str_detect(Subject, "Bio"))
edgesSoc <- edges %>% filter(str_detect(Subject, "Social"))


# Cluster analysis
make_normed_matrix <- function(df, return_ids = FALSE) {
  ids <- df$ID
  df$median_time_per_access <- as.numeric(as_duration_from_string(df$median_time_per_access))
  df$sd_time <- as.numeric(as_duration_from_string(df$sd_time))
  df <- as.data.frame(sapply(df %>% 
                         filter(Role == "S") %>% 
                         select(-ID, -Name, -Subject, -Role, -grade_quartile), 
                       as.numeric))
  #gq <- df$grade_quartile # does not need to be normed
  df <- scale(df)
  if (return_ids) {
    df$ID <- ids
  }
  #df$grade_quartile <- gq
  return(df)
}

nodeMatrixBio <- make_normed_matrix(nodesBio)
nodeMatrixSoc <- make_normed_matrix(nodesSoc)
nodes <- rbind(nodesBio, nodesSoc) %>% filter(Role == "S")
nodeMatrix <- make_normed_matrix(nodes)
dAct <- nodeMatrix[, 2:8]

kmed <- pam(dAct, 5)

nodes <- bind_cols(nodes, data.frame(cluster = kmed$cluster))

# Adding Qual participants
qual <- read_csv(file.path(DATDIR, QUALFILE))
qual <- qual %>% 
  rename(qual = Method) %>% 
  select(-Subject)
nodes <- nodes %>% left_join(qual, by = c("Name"))
nodes$qual_participant <- !is.na(nodes$qual)
nodes[is.na(nodes$qual), ]$qual <- "None"

# Joining cluster and qual back with subject ndoes
nodesBio <- left_join(nodesBio, nodes %>% select(ID, Name, cluster, qual))
nodesSoc <- left_join(nodesSoc, nodes %>% select(ID, Name, cluster, qual))

# Joining grade_quartile, cluster and qual back with aa 
nodeInfo <- nodes %>% 
  select(name = Name, grade_quartile, cluster, qual, qual_participant)
aaBio18 <- left_join(aaBio18, nodeInfo)
aaSoc18 <- left_join(aaSoc18, nodeInfo)

# Making graph structures
networkBio <- tbl_graph(nodes = nodesBio, edges = edgesBio) 
networkSoc <- tbl_graph(nodes = nodesSoc, edges = edgesSoc)

```

# Network analysis on forum posts and replies

Discussion (forum) posts were extraced, and converted into networks. The main algorithms used are:

- _infomap_ for clustering (forming communities)
- _central authority_ for importance
- the force-directed _Kamada Kawai_ algorithm for graph layout

Each *node* in the network represents a person in the course. Each *edge* is the sum of the posts between the people, darker lines mean more posts / replies. Grades are displayed in quartiles, with 4 being the upper quartile and represented by a larger disc.

## Social Psychology

``` {r fig.height = 4}
# Graphs for Social Psychology


gSocAll <- buildGraph(networkSoc)
gSocAll + ggtitle("All Posts for PSY203: Social Psychology") #+ geom_node_text(aes(label = id), size = 2)

gSocNet <- buildTrimGraph(networkSoc)
gSocNet + ggtitle("All Posts for PSY203: Social Psychology (isolated points removed)")

gSocLurker <- buildLurkerGraph(networkSoc %>% activate(nodes) %>% filter(Role == "S"))
gSocLurker + ggtitle("All Posts for PSY203: Social Psychology, with passive forum interactions")

gSocLurkerZoom <- buildLurkerGraph(networkSoc %>% activate(nodes) %>% filter(Role == "S" & !(node_is_isolated())))
gSocLurkerZoom + ggtitle("All Posts for PSY203: Social Psychology, with passive forum interactions (isolated points removed)")

gSocStudents <- buildTrimGraph(networkSoc %>% activate(nodes) %>% filter(Role == "S")) 
gSocStudents + ggtitle("Student only interactions for PSY203: Social Psychology (isolated points removed)")

gSocImportance <- buildImportanceGraph(networkSoc %>% activate(nodes) %>% filter(Role == "S"))
gSocImportance + ggtitle("Student only interaction by Importance for PSY203: Social Psychology (isolated points removed)")
  
gSocCommunity <- buildClusterGraph(networkSoc %>% activate(nodes) %>% filter(Role == "S"))
gSocCommunity + ggtitle("Student only Communities for PSY203: Social Psychology (isolated points removed)")
```

\newpage

## Biopsychology

``` {r fig.height = 4}

gBioAll <- buildGraph(networkBio)
gBioAll + ggtitle("All Posts for PSY208: Biopsychology") #+ geom_node_text(aes(label = id), size = 2)

gBioNet <- buildTrimGraph(networkBio)
gBioNet + ggtitle("All Posts for PSY208: Biopsychology (isolated points removed)")

gBioLurker <- buildLurkerGraph(networkBio %>% activate(nodes) %>% filter(Role == "S"))
gBioLurker + ggtitle("All Posts for PSY208: Biopsychology, with passive forum interactions")

gBioLurkerZoom <- buildLurkerGraph(networkBio %>% activate(nodes) %>% filter(Role == "S" & !(node_is_isolated())))
gBioLurkerZoom + ggtitle("All Posts for PSY208: Biopsychology, with passive forum interactions (isolated points removed)")

gBioStudents <- buildTrimGraph(networkBio %>% activate(nodes) %>% filter(Role == "S")) 
gBioStudents + ggtitle("Student only interactions for PSY208: Biopsychology (isolated points removed)")

gBioImportance <- buildImportanceGraph(networkBio %>% activate(nodes) %>% filter(Role == "S"))
gBioImportance + ggtitle("Student only interaction by Importance for PSY208: Biopsychology (isolated points removed)")
  
gBioCommunity <- buildClusterGraph(networkBio %>% activate(nodes) %>% filter(Role == "S"))
gBioCommunity + ggtitle("Student only Communities for PSY208: Biopsychology (isolated points removed)")
```
