---
title: "Analysis of Engagement in Social Psychology and Biopsychology Courses - Student Activity"
author: "Ben Hicks"
output: html_document
---

The following data is from the Social Psychology (PSY203 and PSY453) and Biopsychology (PSY208 and PSY408) subjects run in 2018 Session 2. The aim of this report is to look at student engagement in these courses through the data available in Blackboard Learn. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F, cache = T)
#knitr::opts_chunk$set(dev = 'png') - - uncomment to reduce file size
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(GGally))
suppressPackageStartupMessages(library(network))
suppressPackageStartupMessages(library(tidygraph))
suppressPackageStartupMessages(library(ggraph))
suppressPackageStartupMessages(library(ggthemes))
suppressPackageStartupMessages(library(lakit))
suppressPackageStartupMessages(library(dendextend))
suppressPackageStartupMessages(library(cluster))
suppressPackageStartupMessages(library(factoextra))
suppressPackageStartupMessages(library(ggsci))
```

``` {r importing, cache=TRUE}
source(file.path('..','R','psych_functions.R'))
DATDIR <- file.path("~", "Data", "Psychology")
BIOAAFILE <- "aa Biopsychology 201860.csv"
SOCAAFILE <- "aa Social Psychology 201860.csv"
MARKSFILE <- "marks Psychology 201860.csv"
BIOCCFILE <- "cc Biopsychology 201860.csv"
SOCCCFILE <- "cc Social Psychology 201860.csv"
POSTSFILE <- "posts Psychology 201860.csv"
QUALFILE <- "qual participants.csv"
ZERODATE <- "2018-05-01" # pre-session date used to colour the site activity structure paths. Affects animation time before student activity

snData <- read_csv(file.path(DATDIR, POSTSFILE), col_types = "cccffn")
# Activity Accumulator Data --------------------------------------
tidy_aa <- function(df) {
  df <- df %>%
    filter(!is.na(id), role == "S") %>%
    mutate(forum_bool = str_detect(tolower(data), "discus|thread|forum"),
                                      content = tidyDataField(data),
                                      name = paste0(lastname, ', ', firstname)) %>%
    rename(session = session_id, id = s_id, login_id = id)
  df <- df %>% mutate(forum_bool = if_else(is.na(forum_bool), FALSE, forum_bool))
  return(df)
} # --------------------------------------------------------------
aaSoc18 <- read_csv(file.path(DATDIR, SOCAAFILE), col_types = "ffTcccccffcccf") # can cause performance hit, if so do not specity columns and add as.factor lines to tidy_aa 
aaSoc18 <- tidy_aa(aaSoc18)

aaBio18 <- read_csv(file.path(DATDIR, BIOAAFILE), col_types = "ffTcccccffcccf")
aaBio18 <- tidy_aa(aaBio18)

# Course contents
tidy_cc <- function(x) {
  x %>% 
    select(pk1, parent_pk1, title, data = main_data) %>% 
    mutate(pk1 = factor(pk1), parent_pk1 = factor(parent_pk1)) %>% 
    mutate(title = str_trunc(title, width = 45))
}
ccSoc18 <- read_csv(file.path(DATDIR, SOCCCFILE)) %>% tidy_cc()
ccBio18 <- read_csv(file.path(DATDIR, BIOCCFILE)) %>% tidy_cc()


# Node preparation (i.e. data for each student)

# Mark data
marks18 <- read_csv(file.path(DATDIR, MARKSFILE)) %>%
  mutate(mark = score / possible, id = factor(id)) 
marksSummary18 <- marks18 %>% 
  group_by(subject_name, student, id) %>%
  summarise(avg_grade = mean(mark, na.rm = T)) %>%
  rename(name = student, subject = subject_name)
marksSoc18 <- marksSummary18 %>% filter(subject == "Social Psychology")
marksBio18 <- marksSummary18 %>% filter(subject == "Biopsychology")

# Discussion views from aa -------------------------------------------
discussion_views <- function(df) {
  df %>%
    select(name, subject, forum_bool, data) %>%
    filter(  forum_bool |
           str_detect(tolower(data), "discus|thread|forum")
         ) %>% 
    group_by(subject, name) %>%
    summarise(ForumViews = n()) %>%
    group_by(name) %>%
    summarise(ForumViews = sum(ForumViews))
} # ------------------------------------------------------------------
discussionViewsBio <- discussion_views(aaBio18)
discussionViewsSoc <- discussion_views(aaSoc18)

# Other interaction data from aa
aaSummaryBio <- aa_summary_by_user(aaBio18)
aaSummaryBio <- merge(aaSummaryBio, unique(aaBio18 %>% select(id, name)))
aaSummarySoc <- aa_summary_by_user(aaSoc18)
aaSummarySoc <- merge(aaSummarySoc, unique(aaSoc18 %>% select(id, name)))


# Node assembly ----------------------------------------------------------
node_merge_and_make <- function(aaData, marksData, discussionsData) {
  df <- aaData %>% select(-id) %>%
  merge(marksData) %>%
  merge(discussionsData, all.x = T) %>%
  select(ID = id, Name = name, Subject = subject, 
         Grade = avg_grade, Accesses = accesses, 
         mean_clicks_per_access, sd_clicks, 
         median_time_per_access, sd_time, 
         TotalTime = total_time,
         ForumViews) %>%
    mutate(grade_quartile = ntile(Grade, 4), ID = factor(ID))
  if (any(is.na(df$ForumViews))) {
    df[is.na(df$ForumViews),]$ForumViews <- 0
  }
  if (any(is.na(df$Accesses))) {
    df[is.na(df$Accesses),]$Accesses <- 0  
  }
  if (any(is.na(df$TotalTime))) {
    df[is.na(df$TotalTime),]$TotalTime <- 0
  }
  df$Role <- "S"
  return(df)
} # ----------------------------------------------------------------------
nodesBio <- node_merge_and_make(aaSummaryBio, marksBio18, discussionViewsBio)
nodesSoc <- node_merge_and_make(aaSummarySoc, marksSoc18, discussionViewsSoc)

# -----------------------------------------------------------------------------
subject_code_to_name <- function(x) {
  ifelse(str_detect(x, "208|458"), "Biopsychology", "Social Psychology")
} # ---------------------------------------------------------------------------

## ADD ##
allStudentIDs <- unique(marks18$id)
allForumIDs <- unique(snData$msgfrom)
teacherIDs <- setdiff(allForumIDs, allStudentIDs)
teacherNodes <- snData %>% 
  filter(msgfrom %in% teacherIDs) %>% 
  select(Name = poster, ID = msgfrom, Subject = subject) %>% 
  mutate(ID = as.factor(ID), 
         Grade = 0, Accesses = 100, ForumViews = 200, Role = "T", Subject = subject_code_to_name(Subject), 
         TotalTime = as.duration(200000), median_time_per_access = as.duration(2000), sd_time = as.duration(200),
         mean_clicks_per_access = 0, sd_clicks = 0, grade_quartile = 0) %>%
  unique()
teacherNodesBio <- teacherNodes %>% filter(Subject == "Biopsychology")
teacherNodesSoc <- teacherNodes %>% filter(Subject == "Social Psychology")

nodesBio <- rbind(nodesBio, teacherNodesBio)
nodesSoc <- rbind(nodesSoc, teacherNodesSoc)

# Edge assembly
edgesAll <- data.frame(from = snData$msgfrom, 
                       to = snData$msgto, 
                       weight = snData$strength, 
                       Subject = snData$subject) %>%
  mutate(Subject = subject_code_to_name(Subject))
questions <- edgesAll %>% 
  filter(to == "NULL") %>% 
  group_by(from, Subject) %>% 
  summarise(Questions = sum(weight))
edges <- edgesAll %>% filter(to != "NULL")
edgesBio <- edges %>% filter(str_detect(Subject, "Bio"))
edgesSoc <- edges %>% filter(str_detect(Subject, "Social"))


# Cluster analysis
make_normed_matrix <- function(df, return_ids = FALSE) {
  ids <- df$ID
  df$median_time_per_access <- as.numeric(as_duration_from_string(df$median_time_per_access))
  df$sd_time <- as.numeric(as_duration_from_string(df$sd_time))
  df <- as.data.frame(sapply(df %>% 
                         filter(Role == "S") %>% 
                         select(-ID, -Name, -Subject, -Role, -grade_quartile), 
                       as.numeric))
  #gq <- df$grade_quartile # does not need to be normed
  df <- scale(df)
  if (return_ids) {
    df$ID <- ids
  }
  #df$grade_quartile <- gq
  return(df)
}

nodeMatrixBio <- make_normed_matrix(nodesBio)
nodeMatrixSoc <- make_normed_matrix(nodesSoc)
nodes <- rbind(nodesBio, nodesSoc) %>% filter(Role == "S")
nodeMatrix <- make_normed_matrix(nodes)
dAct <- nodeMatrix[, 2:8]

kmed <- pam(dAct, 5)

nodes <- bind_cols(nodes, data.frame(cluster = kmed$cluster))

# Adding Qual participants
qual <- read_csv(file.path(DATDIR, QUALFILE))
qual <- qual %>% 
  rename(qual = Method) %>% 
  select(-Subject)
nodes <- nodes %>% left_join(qual, by = c("Name"))
nodes$qual_participant <- !is.na(nodes$qual)
nodes[is.na(nodes$qual), ]$qual <- "None"

# Joining cluster and qual back with subject ndoes
nodesBio <- left_join(nodesBio, nodes %>% select(ID, Name, cluster, qual))
nodesSoc <- left_join(nodesSoc, nodes %>% select(ID, Name, cluster, qual))

# Joining grade_quartile, cluster and qual back with aa 
nodeInfo <- nodes %>% 
  select(name = Name, grade_quartile, cluster, qual, qual_participant)
aaBio18 <- left_join(aaBio18, nodeInfo)
aaSoc18 <- left_join(aaSoc18, nodeInfo)

# Making graph structures
networkBio <- tbl_graph(nodes = nodesBio, edges = edgesBio) 
networkSoc <- tbl_graph(nodes = nodesSoc, edges = edgesSoc)

```

# Activity


``` {r timelines, eval = F}
## Timeline Averages
timelineData <- timelineDataRaw %>% mutate(subject = substr(`subject code`, 3, 8),
                                           session = as.factor(substr(`subject code`, 10, 15)))
timeSum <- timelineData[timelineData$subject != "PSY458",] %>% 
  group_by(week, measure, subject, session) %>%
  summarise(score = sum(score, na.rm = T))

# collecting together different campuses and ommitting some data
nonDiscussion <- c("Avg Site Accesses by Week of Term","Avg Interactions per Week","Avg Minutes per Site Access","Avg Minutes by Day")
timeSumShort <- timelineData[timelineData$subject != "PSY458" &
                              timelineData$measure %in% nonDiscussion,] %>% 
  group_by(week, measure, subject, session) %>%
  summarise(score = sum(score, na.rm = T))
timePlot <- ggplot(timeSumShort, aes(x = week, y = score, color = session)) +
  geom_line() +
  facet_grid(measure ~ subject, scales = "free") +
  theme_tufte() +
  theme(strip.text.y = element_text(size = 8, angle = 0),
        axis.label.y = NULL) +
  scale_color_brewer(palette = "Paired")
timePlot + ggtitle("Average Course Activity by Week of Course")

# t-testing 1st and 3rd graphs for PSY203
t1int <- timeSumShort[timeSumShort$measure == "Avg Interactions per Week" & timeSumShort$subject == "PSY203" & timeSumShort$session == "201760",]$score
t2int <- timeSumShort[timeSumShort$measure == "Avg Interactions per Week" & timeSumShort$subject == "PSY203" & timeSumShort$session == "201860",]$score
t1Min <- timeSumShort[timeSumShort$measure == "Avg Minutes per Site Access" & timeSumShort$subject == "PSY203" & timeSumShort$session == "201760",]$score
t2Min <- timeSumShort[timeSumShort$measure == "Avg Minutes per Site Access" & timeSumShort$subject == "PSY203" & timeSumShort$session == "201860",]$score

ttestInt <- t.test(t2int, t1int, paired = T)
ttestMin <- t.test(t2Min, t1Min, paired = T)

# Old text below:
# The most significant difference between sessions of the results above is for Social Psychology (PSY203) on the *Average Interactions per Week* and *Avg Minutes per Site Access* metrics. The interactions show an increase of between `r round(ttestInt$conf.int[[1]],2)` and `r round(ttestInt$conf.int[[2]], 2)` (95% confidence, paired t-test), and a similar test on the minutes per site access shows an increase of between `r round(ttestMin$conf.int[[1]],2)` and `r round(ttestMin$conf.int[[2]], 2)`. For the latter, it is worth flagging that mean duration on site can be thrown by students leaving the browser open - median would be a better measure here.
```


## Item Activity

We can examine student interactions with the course content at a finer level, and filter according to activities of note.

### Social Psychology

``` {r social_psych_usage, cache = TRUE}
# Using the ACTIVITY_ACCUMULATOR data to examine item activity
socUsage <- aaSoc18 %>%
  select(name, timestamp, content, role, forum_bool) %>%
  filter(role == "S" & timestamp < as.Date('2018-11-01')) %>%
  mutate(Activity = if_else(
    str_detect(content, "Week 2: The Social Self|Week 3:Perc|Week 4:Stereo|Week 5:Att|Week 6:Conf|Week 10:Att|Week 11:Help|Applied Social"), 
    "Tutorial",
    if_else(forum_bool, "Forum", "Other")
    )
    )%>%
  mutate(Activity = if_else(is.na(Activity), "Other", Activity)) %>%
  rename(Name = name) %>% 
  left_join(nodesSoc %>% 
          select(Name, Accesses, TotalTime, grade_quartile, ForumViews) %>% 
          unique(), by = c("Name"))

socGameActPlot <- socUsage %>% 
  ggplot(aes(x = timestamp, y = grade_quartile, color = ForumViews)) +
  scale_color_gradient2(name = "Forum Participation", 
                         low = "Black",
                          mid = "Dark Blue",
                         high = "Red",
                         midpoint = 3.5,
                         trans = "log1p",
                         breaks = c(2, 10, 50, 250, 1250)) +  
  geom_jitter(alpha = 0.01) +
  scale_y_continuous(name = "Grade Quartile") +
  ggthemes::theme_tufte() + 
  theme(axis.title.x = element_blank()) + 
  facet_grid(Activity ~ .) +
  ggtitle("Usage of items in Social Psychology")

socUsage %>% group_by(Activity) %>% tally() %>% knitr::kable()

socGameActPlot

```

``` {r site_nav_functions, eval = T}
# Looking at how students navigated the course
edges_from_aa <- function(x) {
  aa <- x %>% 
    filter(role == 'S', !is.na(content_pk1), timestamp < as.POSIXct('2018-10-30')) %>% 
    select(id, time = timestamp, content_pk1, session, 
           grade_quartile, cluster, qual) %>% 
    mutate(id = factor(id), content_pk1 = factor(content_pk1)) %>% 
    arrange(id, time)
  aa_edges <- tibble(from = aa$content_pk1[1:(nrow(aa) - 1)],
                   to =   aa$content_pk1[2:nrow(aa)],
                   time = aa$time[2:nrow(aa)],
                   student_from = aa$id[1:(nrow(aa) - 1)],
                   student = aa$id[2:nrow(aa)],
                   session_from = aa$session[1:(nrow(aa) - 1)],
                   session = aa$session[2:nrow(aa)],
                   grade_quartile = aa$grade_quartile[1:(nrow(aa) - 1)],
                   cluster = aa$cluster[1:(nrow(aa) - 1)],
                   qual = aa$qual[1:(nrow(aa) - 1)])
  aa_edges <- aa_edges %>% 
    filter(student == student_from, session == session_from) %>% 
    select(from, to, time, student, grade_quartile, cluster, qual) %>% 
    mutate(weight = 2)
  return(aa_edges)
}

edges_from_cc <- function(x) {
  cc_edges <- x %>% 
    select(from = parent_pk1, to = pk1) %>% 
    filter(!is.na(from)) %>% 
    mutate(weight = 1, time = as.POSIXct(ZERODATE)) # used manual set time - needs to be well before other data
  return(cc_edges)
}

nodes_from_cc <- function(x, subject) {
  x <- x %>% select(id = pk1, title, data) %>% 
    mutate(forum_bool = str_detect(tolower(data), "discus|thread|forum")) %>% 
    mutate(forum_bool = if_else(is.na(forum_bool), FALSE, forum_bool))
  if (str_detect(subject, "Bio|bio")) {
    x <- x %>% 
      mutate(PageType = if_else(str_detect(title, "lab|Lab"), "Lab", 
                                if_else(forum_bool, "Forum", "Other"))) %>%
      mutate(PageType = if_else(is.na(PageType), "Other", PageType)) %>% 
      mutate(PageType = factor(PageType, levels = c("Forum", "Lab", "Other")))
  }
  if (str_detect(subject, "Soc|soc")) {
   x <- x %>% 
     mutate(PageType = if_else(str_detect(title, "Week 2: The Social Self|Week 3:Perc|Week 4:Stereo|Week 5:Att|Week 6:Conf|Week 10:Att|Week 11:Help|Applied Social"), 
                               "Tutorial",
                               if_else(forum_bool, "Forum", "Other"))) %>%
     mutate(PageType = if_else(is.na(PageType), "Other", PageType)) %>% 
     mutate(PageType = factor(PageType, levels = c("Forum", "Tutorial", "Other")))
   }
  return(x %>% select(-data))
  }

nodes_from_aa <- function(x) {
  x %>% 
    filter(!is.na(content_pk1)) %>% 
    left_join(nodes %>% select(id = ID, ))
    select(id = content_pk1, cluster, qual) %>% # might need to calc grade quartile
    unique()
}

build_site_activity_graph_list <- function(cc_nodes, cc_edges, aa_nodes, aa_edges) {
  cc_graph <- tbl_graph(nodes = cc_nodes, edges = cc_edges)
  cc_layout <- create_layout(cc_graph, layout = "nicely")
  np <- data.frame(x = cc_layout$x, y = cc_layout$y)
  aa_graph <- tbl_graph(nodes = cc_nodes, edges = aa_edges)
  joined_graph <- cc_graph %>% graph_join(aa_graph) 
  joined_graph <- joined_graph %>% 
    activate(edges) %>% 
    mutate(student = if_else(is.na(student), "_course structure", as.character(student)))
  return(list(joined_graph, np))
}


draw_site_graph <- function(graph_list){
  graph <- graph_list[[1]] %>% 
    activate(edges) %>% 
    filter(weight == 1)
  np <- graph_list[[2]]
  g <- ggraph(graph, layout = "manual", node.positions = np) +
    geom_node_point(aes(color = PageType)) + 
    geom_node_text(aes(label = title), alpha = 0.4, repel = T, size = 2) +
    geom_edge_link0() +
    theme_graph() +
    theme(#legend.position = "none",
        text = element_text(family = "TT Times New Roman"))
  return(g)
}

draw_site_activity_graph <- function(graph_list, alpha_low = 0.1) {
  g <- ggraph(graph_list[[1]], layout = "manual", node.positions = graph_list[[2]]) +
    geom_node_point() + 
    geom_edge_link0(aes(color = as.numeric(time), 
                      width = weight, 
                      alpha = 2 - weight,
                      group = time)) +
    scale_edge_alpha_continuous(range = c(alpha_low, 1)) +
    scale_edge_width(range = c(0.6,1)) +
    scale_edge_color_gradientn(colors = c("black", "#47ADED", "#D81C41")) +
    theme_graph() +
    theme(legend.position = "none",
        text = element_text(family = "TT Times New Roman"))
  return(g)
}
```

#### Site structure and activity

``` {r soc_site, cache = T}
socSiteGraphList <- build_site_activity_graph_list(
  cc_nodes = nodes_from_cc(ccSoc18, subject = "Soc"),
  cc_edges = edges_from_cc(ccSoc18),
  aa_nodes = nodes_from_aa(aaSoc18),
  aa_edges = edges_from_aa(aaSoc18))

draw_site_graph(socSiteGraphList) + ggtitle("Social Psychology site structure")
```

``` {r soc_site_nav, cache = TRUE}
draw_site_activity_graph(socSiteGraphList, alpha_low = 0.002) + ggtitle("Social Psychology activity")
```

\newpage

### Biopsychology Activity

``` {r biopsych_usage, cache = TRUE}
# Using the ACTIVITY_ACCUMULATOR data to examine item activity

bioUsage <- aaBio18 %>%
  select(name, timestamp, content, role, forum_bool) %>%
  filter(role == "S" ) %>%
  #& timestamp < as.Date('2018-11-01')) %>%
  mutate(Activity = if_else(str_detect(tolower(content), "game"), "Game",
                                   if_else(str_detect(tolower(content), "in the lab"), "In the lab",
                                           if_else(forum_bool, "Forum", "Other")))) %>%
  mutate(Activity = if_else(is.na(Activity), "Other", Activity)) %>%
  rename(Name = name) %>%
  merge(nodesBio %>% select(Name, Accesses, grade_quartile, ForumViews, 
                            mean_clicks_per_access, sd_clicks, median_time_per_access, sd_time,
                            TotalTime), all.x = T)



bioGameActPlot <- bioUsage %>% 
  ggplot(aes(x = timestamp, y = grade_quartile, color = ForumViews)) +
  scale_color_gradient2(name = "Forum Participation", 
                         low = "Black",
                          mid = "Dark Blue",
                         high = "Red",
                         midpoint = 3.5,
                         trans = "log1p",
                         breaks = c(2, 10, 50, 250, 1250)) +  
  geom_jitter(alpha = 0.002) +
  scale_y_continuous(name = "Grade Quartile") +
  ggthemes::theme_tufte() + 
  theme(axis.title.x = element_blank()) + 
  facet_grid(Activity ~ .) +
  ggtitle("Usage of items in Biopsychology")

bioUsage %>% group_by(Activity) %>% tally() %>% knitr::kable()

bioGameActPlot

bioGameActPlotFocus <- bioUsage %>% filter(timestamp < as.Date('2018-10-15')) %>%
  ggplot(aes(x = timestamp, y = grade_quartile, color = ForumViews)) +
  scale_color_gradient2(name = "Forum Participation", 
                         low = "Black",
                          mid = "Dark Blue",
                         high = "Red",
                         midpoint = 3.5,
                         trans = "log1p",
                         breaks = c(2, 10, 50, 250, 1250)) +  
  geom_jitter(alpha = 0.002) +
  scale_y_continuous(name = "Grade Quartile") +
  ggthemes::theme_tufte() + 
  theme(axis.title.x = element_blank()) + 
  facet_grid(Activity ~ .) +
  ggtitle("Usage of items in Biopsychology - focus on active session")


bioGameActPlotFocus
```

#### Site structure and activity

``` {r bio_site_structure}
bioSiteGraphList <- build_site_activity_graph_list(
  cc_nodes = nodes_from_cc(ccBio18, subject = "Bio"),
  cc_edges = edges_from_cc(ccBio18),
  aa_nodes = nodes_from_aa(aaBio18),
  aa_edges = edges_from_aa(aaBio18))
```

``` {r bio_site, cache = T, eval = T}
draw_site_graph(bioSiteGraphList) + ggtitle("Biopsychology site structure")
```

``` {r bio_site_nav, cache = T, eval = T}
draw_site_activity_graph(bioSiteGraphList, alpha_low = 0.002) + ggtitle("Biopsychology activity")
```