---
title: "Analysis of Engagement in Social Psychology and Biopsychology Courses - Clustering"
author: "Ben Hicks"
output: html_document
---

The following data is from the Social Psychology (PSY203 and PSY453) and Biopsychology (PSY208 and PSY408) subjects run in 2018 Session 2. The aim of this report is to look at student engagement in these courses through the data available in Blackboard Learn. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
#knitr::opts_chunk$set(dev = 'png') - - uncomment to reduce file size
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(GGally))
suppressPackageStartupMessages(library(network))
suppressPackageStartupMessages(library(tidygraph))
suppressPackageStartupMessages(library(ggraph))
suppressPackageStartupMessages(library(ggthemes))
suppressPackageStartupMessages(library(lakit))
suppressPackageStartupMessages(library(dendextend))
suppressPackageStartupMessages(library(cluster))
suppressPackageStartupMessages(library(factoextra))
suppressPackageStartupMessages(library(ggsci))
```

``` {r importing, cache=TRUE}
source(file.path('..', 'R', 'psych_functions.R'))
source(file.path('..', 'R', 'import-data.R'))
```


# Modelling

Clustering seemed to be most parsimonious choosing the *k-medoids* algorithm with 5 clusters across both subjects, and *excluding academic data*. See the document *cluster analysis psych* for more detail on why this was chosen.

## Principal Component Analysis

Using the 

## Cluster Analysis


``` {r modelling_setup}
# feed in grouped_by df
cluster_summary_table <- function(x) {
  x %>%
    summarise(n = n(), 
              Biopsych = sum(Subject == "Biopsychology"),
              SocPsych = sum(Subject == "Social Psychology"),
              Grade = mean(Grade, na.rm = T),
              Grade_sd = sd(Grade, na.rm = T),
              HDs = sum(Grade > 80, na.rm = T),
              Accesses = mean(Accesses, na.rm = T),
              clicks_pa = mean(mean_clicks_per_access, na.rm = T),
              sd_clicks = mean(sd_clicks, na.rm = T),
              median_time_pa = as.duration(mean(median_time_per_access)),
              sd_time = as.duration(mean(sd_time)),
              TotalTime = as.duration(mean(TotalTime)),
              ForumViews = mean(ForumViews)) %>% 
    knitr::kable()
}

fviz_cluster(kmed)

nodes %>% group_by(cluster) %>% cluster_summary_table()

```

## Activity by Cluster

### Social Psychology

``` {r social_psych_usage, cache = TRUE}
# Using the ACTIVITY_ACCUMULATOR data to examine item activity
socUsage <- aaSoc18 %>%
  select(name, timestamp, content, role, forum_bool) %>%
  filter(role == "S" & timestamp < as.Date('2018-11-01')) %>%
  mutate(Activity = if_else(
    str_detect(content, "Week 2: The Social Self|Week 3:Perc|Week 4:Stereo|Week 5:Att|Week 6:Conf|Week 10:Att|Week 11:Help|Applied Social"), 
    "Tutorial",
    if_else(forum_bool, "Forum", "Other")
    )
    )%>%
  mutate(Activity = if_else(is.na(Activity), "Other", Activity)) %>%
  rename(Name = name) %>% 
  left_join(nodes %>% filter(Subject == "Social Psychology") %>% 
          select(Name, Accesses, TotalTime, grade_quartile, cluster) %>% 
          unique(), by = c("Name")) %>% 
  mutate(cluster = factor(cluster))

socGameActPlot <- socUsage %>% 
  ggplot(aes(x = timestamp, y = grade_quartile, color = cluster)) +
  geom_jitter(alpha = 0.015) +
  scale_y_continuous(name = "Grade Quartile") +
  ggthemes::theme_tufte() + 
  theme(axis.title.x = element_blank()) + 
  facet_grid(Activity ~ .) +
  ggtitle("Usage of items in Social Psychology")

socGameActPlot
```




``` {r soc_site_nav, cache = T}
socSiteGraphList <- build_site_activity_graph_list(
  cc_nodes = nodes_from_cc(ccSoc18, subject = "Soc"),
  cc_edges = edges_from_cc(ccSoc18),
  aa_nodes = nodes_from_aa(aaSoc18),
  aa_edges = edges_from_aa(aaSoc18))

draw_site_graph(socSiteGraphList) + ggtitle("Social Psychology site structure")

draw_site_activity_graph(socSiteGraphList, alpha_low = 0.02) + facet_wrap(~ cluster) + ggtitle("Social Psychology activity by cluster")
```

\newpage

### Biopsychology

``` {r biopsych_usage, cache = TRUE}
# Using the ACTIVITY_ACCUMULATOR data to examine item activity

bioUsage <- aaBio18 %>%
  select(name, timestamp, content, role, forum_bool) %>%
  filter(role == "S" ) %>%
  #& timestamp < as.Date('2018-11-01')) %>%
  mutate(Activity = if_else(str_detect(tolower(content), "game"), "Game",
                                   if_else(str_detect(tolower(content), "in the lab"), "In the lab",
                                           if_else(forum_bool, "Forum", "Other")))) %>%
  mutate(Activity = if_else(is.na(Activity), "Other", Activity)) %>%
  rename(Name = name) %>%
  merge(nodes %>% filter(Subject == "Biopsychology") %>%  
          select(Name, Accesses, grade_quartile, cluster, 
                            mean_clicks_per_access, sd_clicks, median_time_per_access, sd_time,
                            TotalTime), all.x = T) %>% 
  mutate(cluster = factor(cluster))

bioGameActPlotFocus <- bioUsage %>% filter(timestamp < as.Date('2018-10-15')) %>%
  ggplot(aes(x = timestamp, y = grade_quartile, color = cluster)) +
  geom_jitter(alpha = 0.01) +
  scale_y_continuous(name = "Grade Quartile") +
  ggthemes::theme_tufte() + 
  theme(axis.title.x = element_blank()) + 
  facet_grid(Activity ~ .) +
  ggtitle("Usage of items in Biopsychology with clusters")

bioGameActPlotFocus
```

``` {r bio_site_nav, cache = T}
bioSiteGraphList <- build_site_activity_graph_list(
  cc_nodes = nodes_from_cc(ccBio18, subject = "Bio"),
  cc_edges = edges_from_cc(ccBio18),
  aa_nodes = nodes_from_aa(aaBio18),
  aa_edges = edges_from_aa(aaBio18))

draw_site_graph(bioSiteGraphList) + ggtitle("Biopsychology site structure")

draw_site_activity_graph(bioSiteGraphList, alpha_low = 0.02) + facet_wrap(~ cluster) + ggtitle("Biopsychology activity by cluster")
```

\newpage

## Activity profile of clusters

In progress - spectral analysis of clusters

``` {r spectral_analysis_of_timestamps, eval = F}
bioStampsAndClusters <- inner_join(nodes%>% 
                                     select(name = Name, cluster),
                                aaBio18 %>% 
                                  select(name, timestamp, session))

c1sample <- sample(bioStampsAndClusters[bioStampsAndClusters$cluster == 1,]$name, 5)
c2sample <- sample(bioStampsAndClusters[bioStampsAndClusters$cluster == 2,]$name, 5)
c3sample <- sample(bioStampsAndClusters[bioStampsAndClusters$cluster == 3,]$name, 5)
c4sample <- sample(bioStampsAndClusters[bioStampsAndClusters$cluster == 4,]$name, 5)
c5sample <- sample(bioStampsAndClusters[bioStampsAndClusters$cluster == 5,]$name, 5)

map(c1sample, function(x)timelist_to_difference(bioStampsAndClusters %>% filter(name == x)))

c11 <- data.frame(timelist_to_difference(bioStampsAndClusters[bioStampsAndClusters$name == c1sample[[1]], ]))
plot_timestamp_spectrum(tl, group = id) + facet_wrap(~ id)
```

In progress - animated activity by cluster

``` {r animated_activity_of_clusters, eval = F}
```

\newpage

## Network analysis on forum posts and replies, highlighting clusters

### Social Psychology

``` {r fig.height = 4}
# Graphs for Social Psychology 
gSocCluster <- buildClusterKMetGraph(networkSoc %>% 
                                       activate(nodes) %>% 
                                       filter(Role == "S"))
gSocCluster + ggtitle("Clusters for PSY203: Social Psychology")
```

\newpage

### Biopsychology

``` {r fig.height = 4}
gBioCluster <- buildClusterKMetGraph(networkBio %>% 
                                       activate(nodes) %>% 
                                       filter(Role == "S"))
gBioCluster + ggtitle("Clusters for PSY208: Biopsychology")
```

